### 2018-11-28 paper.tex
update 11/28 in Tokyo

\documentclass[runningheads]{llncs}

\usepackage{color}

\usepackage{graphicx}
\usepackage{tikz}
\usepackage{calc}

\usepackage{amsmath, amssymb}
\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{wasysym}

\usetikzlibrary{automata}

\title{On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}
}
\author{
Szil\'{a}rd Zsolt Fazekas\inst{1} \and
Shinnosuke Seki\inst{2}\thanks{Corresponding author}}
\institute{
Akita University, 
Graduate School of Engineering Science, 
1-1 Tegate Gakuen-machi, Akita, 0108502, Japan \\
\email{szilard.fazekas@ie.akita-u.ac.jp}
\and
The University of Electro-Communications, 
Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \\
\email{s.seki@uec.ac.jp}
}

\begin{document}

\maketitle

\begin{abstract}
We investigate simple oritatami systems in an attempt to establish lower bounds on the size and complexity of computationally universal systems. 
In particular, we look at oritatami systems, where the folding sequence consists of a number of beads of the same type and show that under reasonable assumptions, these systems are not universal.
\end{abstract}

%------------------------------------------------------
	\section{Introduction}
%------------------------------------------------------

Transcription is the first essential step of gene expression, when a DNA template sequence is copied into a complementary single stranded RNA sequence by a `copy machine' called RNA polymerase. 
The RNA strand is synthesized letter by letter according to the complimentarity relation ${\tt A} \rightarrow {\tt U}$, ${\tt G} \rightarrow {\tt C}$, ${\tt C} \rightarrow {\tt G}$, and ${\tt T} \rightarrow {\tt A}$ and folds up during a process called co-transcriptional folding.

In a recent breakthrough in molecular engineering by Geary, Rothemund and Andersen~\cite{GearyRothemundAndersen2014} the co-transcriptional folding of RNA is controlled by careful design of the DNA template. As demonstrated in laboratory, this method, called RNA Origami, makes it possible to build rectangles out of RNA strands. 
Geary et al.~\cite{GeMeScSe2016} proposed a mathematical model for this process, called oritatami system.
It has been just shown in \cite{GeMeScSe2018} that the model is efficiently Turing universal by simulating cyclic tag systems introduced by Cook~\cite{Cook2004}. 
The simulation involves a very large and complex oritatami system. 
One future direction of research is to find smaller universal oritatami system.

Closely related is the question of where not to look for universal systems, i.e., what are the limitations of simple oritatami system. 
In search for simple oritatami systems, there are a number of restrictions one can pose on them:
\begin{itemize}
\item bounds on the delay, the number of bead types or the arity;
\item bounds on the size of the primary structure or of the attraction rule set;
\item structural conditions on the primary structure or the attraction rule set.
\end{itemize}

%Rogers and Seki~\cite{SimpleSim} showed that certain OS with small delays are not universal.

%Han, Kim, Rogers and Seki\cite{SelfRemove} proved that it is possible to avoid self-attraction rules without changing the behavior of OS.

In this paper we start a new line of study which concerns the alphabet size for the primary structure, i.e., the number of bead types.

\begin{figure}[tb]
\begin{tikzpicture}

\node (origin) at (0, 0) {};

\draw[-latex] (origin) -- ++(0:3) node[below] {$\delta$}; 
\node at (1, -0.2) {1};
\node at (2, -0.2) {2};
\draw[-latex] (origin) -- ++(90:5) node[left] {$\alpha$}; 
\node at (-0.2, 1) {1};
\node at (-0.2, 2) {2};
\node at (-0.2, 3) {3};
\node at (-0.2, 4) {4};

%Results
\node at (1, 4) {$O(n^2)$};
\node at (1, 3) {$O(n)$};
\node at (1, 2) {$O(n^2)$};
\node at (1, 1) {$O(n)$ \cite{DemaineDNA24}};

\node at (2, 1) {$O(n^2)$};

\end{tikzpicture}
\caption{Summary of the results.}
\label{fig:summary}
\end{figure}

%------------------------------------------------------
	\section{Preliminaries}
%------------------------------------------------------
\input{uos_preliminaries.tex}

\begin{proposition}\label{prop:check_validity}
	For any rule set $R$, arity $\alpha$ and conformation $C = (P,w,H)$ it is possible to check whether $C$ is $R$-valid and whether $C\in \mathcal{C}_{\leq \alpha}$ in time $\mathcal{O}(|H|\cdot|w|\cdot|R|)$.
\end{proposition}
\begin{proof}
	To check whether $C$ is $R$-valid:
	\begin{enumerate}
		\item FOR each $(i,j)\in H$:
		\item \hspace{1cm} IF $(w[i],w[j])\notin R$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|w|\cdot|R|)$ time for any reasonable representation of $w$ and $R$, hence the whole process takes $\mathcal{O}(|H|\cdot |w|\cdot|R|)$ time.	
	To check the arity constraint $C\in \mathcal{C}_{\leq \alpha}$: 
	\begin{enumerate}
		\item FOR each $i\in \{1,\dots,|w|\}$:
		\item \hspace{1cm} IF $\mathrm{degree}(i)=|\{j | (i,j)\in H \}|>\alpha$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|H|)$ time for any reasonable representation of $H$, hence the whole process takes $\mathcal{O}(|w|\cdot|H|)$ time.
\end{proof}
\begin{theorem}\label{thm:OS_to_2dTM}
	There is an algorithm that simulates any oritatami system $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ in time $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$. 
\end{theorem}
\begin{proof}
	Take any step in the computation, up to which some $i \ge 0$ first beads of $w$ have been stabilized, with the last bead at a point $p$. 
	The number of all possible elongations of the current conformation by the next $\delta$-beads is $(6 \times 5^{\delta-1}) \times ((2^4)^{\delta-1} \times 2^5) \in 2^{O(\delta)}$. 
	By Proposition~\ref{prop:check_validity}, we can check for each of these elongations whether its arity is at most $\alpha$ or not and whether it is $R$-valid or not in time $\mathcal{O}((2^4)^{\delta-1}\cdot2^5\cdot \delta\cdot|R|)=2^{\mathcal{O}(\delta)}\cdot|R|$.  Therefore, the total running time is $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$.
\end{proof}

\begin{corollary}
	For fixed $\delta$ and $\alpha$, the class of problems solvable by oritatami systems $(\Sigma, R, \delta, \alpha, \sigma, w)$ is included in $\mathrm{DTIME}(n^3)$.
\end{corollary}
\begin{proof}
	The claim follows from Theorem~\ref{thm:OS_to_2dTM} and the fact that $|R|$ is implicitly bounded by $|w|^2$.
\end{proof}

%\begin{corollary}\label{polytranscript}
%	Let $k$ be a non-negative number. Let $\Sigma$, $H$, $\delta$, $\alpha$ be fixed and the transcript length bounded polynomially by the seed length, i.e., $|w|\in O(|\sigma|^k)$. Then, the language of seeds $\sigma$ accepted by an OS $(\Sigma, R, \delta, \alpha, \sigma, w)$ is in $\mathrm{DTIME}(|\sigma|^{2k})$.
%\end{corollary}

%If the transcript length is polynomially bounded by the seed length, than the accepted language is in $\mathrm{P}$.

Because of the time hierarchy theorems, we know that $\mathrm{P}\subsetneq \mathrm{EXP}$, so we can conclude that OS which cannot deterministically fold transcripts of length exponential in the length of the seed are not computationally universal.

%-------------------------------------------------------
	\section{Problem description}
%-------------------------------------------------------

In \cite{DHOPRSST2018}, Demaine et al. proved that at delay 1 and arity 1, an oritatami system can fold upon the seed of size $n$ only the first $9n$ beads of a transcript deterministically. 
We consider this finiteness problem for unary oritatami systems under various settings of the values of delay and arity, which is formalized as follows.

\begin{problem}\label{prob:det_unary_length}
Give an upper bound on the length of a transcript of a delay-$\delta$, arity-$\alpha$ deterministic unary oritatami system whose seed is of length $n$ by a function in $\delta$, $\alpha$, and $n$. 
\end{problem}

Results on this problem are summarized in Fig.~\ref{fig:summary}.

\section{Case of $\delta = 1$}

%//////////////////////////////////////////////////////%
%//////////////////////////////////////////////////////%
%Kohei
%//////////////////////////////////////////////////////%
%//////////////////////////////////////////////////////%l.177 - 1312

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \fill (0,0) circle[radius = 0.1];
      \fill (1.5,0) circle[radius = 0.1];
      \fill (3,0) circle[radius = 0.1];
      \fill (4.5,0) circle[radius = 0.1];
      \fill (6,0) circle[radius = 0.1];

      \draw[dashed] (0,0)--(1.5,0);
      \draw[dashed] (1.5,0)--(3,0);
      \draw[dashed] (3,0)--(4.5,0);
      \draw[dashed] (4.5,0)--(6,0);
      \draw[->] (0,0)--(60:0.75);
      \draw[->] (1.5,0)--++(60:0.75);
      \draw[->] (3,0)--++(60:0.75);
      \draw[->] (4.5,0)--++(60:0.75);
      \draw (0,0)--(60:1.5);
      \draw (1.5,0)--++(60:1.5);
      \draw (3,0)--++(60:1.5);
      \draw (4.5,0)--++(60:1.5);

      \begin{scope}[shift=(60:1.5)]
        \fill (0,0) circle[radius = 0.1];
        \fill (1.5,0) circle[radius = 0.1];
        \fill (3,0) circle[radius = 0.1];
        \fill (4.5,0) circle[radius = 0.1];
        \draw[dashed] (0,0)--(1.5,0);
        \draw[dashed] (1.5,0)--(3,0);
        \draw[dashed] (3,0)--(4.5,0);
        \draw[->] (0,0)--(-60:0.75);
        \draw[->] (1.5,0)--++(-60:0.75);
        \draw[->] (3,0)--++(-60:0.75);
        \draw[->] (4.5,0)--++(-60:0.75);

        \draw (0,0)--(-60:1.5);
        \draw (1.5,0)--++(-60:1.5);
        \draw (3,0)--++(-60:1.5);
        \draw (4.5,0)--++(-60:1.5);
      \end{scope}

    \end{tikzpicture}
    \caption{zig-zag conformation}
    \label{TTT_zigzag}
  \end{center}
\end{figure}

\subsection{Introduction}

In this section, we consider on finiteness of structures produced deterministically at delay 1. In result, cases of arity 1 and 3 can only yield finite structures which is size of $\mathcal{O} (n)$, and cases of arity 4 and more can only yield finite structures which is size of $mathcal{O} (n^2)$, and a case of arity 2 can yield infinite structures but they are only the zig-zag conformation shown in Fig.\ref{TTT_zigzag}.

%In this section, we prove that unary oritatami system can form infinitely at delay 1 and arity 2 deterministically and moreover that the only infinite conformations which its oritatami system can yield is only the zig-zag conformation shown in Fig.\ref{TTT_zigzag}.

Let $\Xi$ be a deterministic oritatami system of delay 1 and arity 2. Assume its seed $\sigma$ consists of $n$ beads.
%Let us denote its transcript by $w = w_1 w_2 w_3 \cdots$ for some $w_1,w_2,w_3 \in \Sigma = \{e\}$.
For $i \geq 0$ let $C_i$ be the unique elongation of $\sigma$ by $w[1..i]$, that is, foldable by $\Xi$. Hence $C_0 = \sigma$.

Let us consider the stabilization of the $i$-th bead $a_i$ upon $C_{i-1}$. The bead cannot collaborate with any succeeding bead $w[i+1],w[i+2],\cdots$ at delay 1. There are just two ways to get stabilized at delay 1. One way is to be bound to another bead. The other way is through a $tunnel\ section$. A tunnel section consists of four beads that occupy four neighbors of a point (Fig.\ref{TTT_tunnel_intro}). 
Accordingly, how they are stabilized can be described by a binary sequence $S$ of $b$'s (bound) and $t$'s (tunnel section); priority is given to $t$, that is, $S[i] = t$ if the $i$-th bead $w_i$ is stabilized not only by being bound but also through a tunnel section.

Assume that four of the six neighbors of a point $p$ are occupied by beads $a_{j_1},a_{j_2},a_{j_3},a_{j_4}$ while the other two are free. We call such a bead $p$ as $inside\ of\ a\ tunnel$ and such beads $p^\prime$ as $entrance\ of\ a\ tunnel$ without a case that $p^\prime$ is inside of a tunnel. If the beads $w[i-2]$ and $w[i-1]$ are stabilized respectively at one of the two free neighbors and at $p$ one after another, then the next bead $w[i]$ cannot help but be stabilized at the other free neighbor. In this way, $w[i]$ can get stabilized without being bound.

%If a bead is stabilized through a tunnel section, then it can provide some bonds. Let us consider bonds of $C_i$. $C_i$ is represented $C = (W,P,H)$ where $|W| = i + n$ $C_i$ contains $\alpha \cdot (i + n) - 2|H|$ bonds because $C_i$ consists of $i + n$ beads and a bead has just $\alpha$ bonds and then $2 |H|$ of the those bonds are already used. However, even if a bead has an available bond, $w[j]$ $(j > i)$ might not be able to use this bond because the bond has possibility that it is blocked by transcripts $w[i+1..j-1]$. Number of $binding\ capabilities$ does not contain that case so that it is at most $\alpha \cdot (i + n) - 2|H|$.

We say that point $p$ is reachable from a conformation $C$ if there exists a directed path $P^\prime$ from the last point of $C$ that does not cross the path of $C$. We define $binding\ capability$ with reachable.

\begin{definition}[binding capability]
Let $B_i$ be $( \{ (h, i) | {}^\forall h < i \} \cup \{ (i, j) | {}^\forall j > i \}) \cap H$. Moreover, let $R_i$ be a set of neighbors of $w[i]$ that are free and reachable from $C_j$ where $C_j$ is a conformation which stabilized until $w[j]$.
 We represent the number of binding capabilities of a conformation $C_j$ as $\#bc(C_j)$. $\#bc(C_j)$ is defined by $\sum^{j}_{k=-n+1} \min \{ |B_k|, |R_k|  \}$.
\end{definition}

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw (0:0) circle [radius=0.05];
      \draw (0:1) circle [radius=0.05];
      \draw (180:1) circle [radius=0.05];
      \node[above] at (0:1) { $p^\prime$ };
      \node[above] at (0:0) { $p$ };
      \node[above] at (180:1) { $p^\prime$ };

      \fill (60 : 1) circle [radius=0.1];
      \fill (-60 : 1) circle [radius=0.1];
      \fill (120 : 1) circle [radius=0.1];
      \fill (-120 : 1) circle [radius=0.1];
      \node[above] at (60 :1) { $a_{j_1}$ };
      \node[above] at (-60 :1) { $a_{j_2}$ };
      \node[above] at (120 :1) { $a_{j_3}$ };
      \node[above] at (-120 :1) { $a_{j_4}$ };

      \begin{scope}[shift=(0:5)]
        \fill[blue] (0:0) circle [radius=0.1];
        \draw (0:1) circle [radius=0.05];
        \fill (180:1) circle [radius=0.1];
        \node[above] at (180:1.2) { $a_{i-2}$ };
        \node[above] at (0:0) { $a_{i-1}$ };
        \draw[->, blue] (180:0.9) -- (180:0.1);

        \fill (60 : 1) circle [radius=0.1];
        \fill (-60 : 1) circle [radius=0.1];
        \fill (120 : 1) circle [radius=0.1];
        \fill (-120 : 1) circle [radius=0.1];
        \node[above] at (60 :1) { $a_{j_1}$ };
        \node[above] at (-60 :1) { $a_{j_2}$ };
        \node[above] at (120 :1) { $a_{j_3}$ };
        \node[above] at (-120 :1) { $a_{j_4}$ };
      \end{scope}
      \begin{scope}[shift=(0:10)]
        \fill (0:0) circle [radius=0.1];
        \fill[blue] (0:1) circle [radius=0.1];
        \fill (180:1) circle [radius=0.1];
        \node[above] at (180:1) { $a_{i-2}$ };
        \node[above] at (0:0) { $a_{i-1}$ };
        \node[above] at (0:1) { $a_i$ };
        \draw[->] (180:0.9) -- (180:0.1);
        \draw[->,blue] (0:0.1) -- (0:0.9);

        \fill (60 : 1) circle [radius=0.1];
        \fill (-60 : 1) circle [radius=0.1];
        \fill (120 : 1) circle [radius=0.1];
        \fill (-120 : 1) circle [radius=0.1];
        \node[above] at (60 :1) { $a_{j_1}$ };
        \node[above] at (-60 :1) { $a_{j_2}$ };
        \node[above] at (120 :1) { $a_{j_3}$ };
        \node[above] at (-120 :1) { $a_{j_4}$ };
      \end{scope}
    \end{tikzpicture} 
    \caption{Through a tunnel section}
    \label{TTT_tunnel_intro}
  \end{center}
\end{figure}

\begin{theorem}[Tunnel Troll Theorem]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 2$. If there are indices $i$ and $j$ such that $S[i..j+1] = bbt^{(j-i-1)}b$, then $\#bc(C_{i-1}) > \#bc(C_j)$ and if $S[i..j+1] = bt^lbt^mb \quad (l + m = j-i-1)$, then $\#bc(C_{i-1}) > \#bc(C_j)$. On the other hand, at $\delta = 1$ and $\alpha \geq 3$, if S$[k] = t$, then $\#bc(C_{k-1}) > \#bc(C_k)$.
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\proof%{Proof of Tunnel Troll Theorem}
Assume $\Xi$ is deterministic.
Each bead in the transcript is bound either inside a tunnel or outside. If a bead is stabilized inside a tunnel, then its successor is already decided the position either inside of a tunnel or outside. Moreover, If a bead is stabilized outside a tunnel, then its position is an entrance of a tunnel or otherwise.

Tunnel sections have three possible shapes up to symmetry : straight($A$), obtuse($B$) and acute($C$) turn (Fig. \ref{TTT_tunnel}), and we will consider each of those.

\begin{lemma}
\label{TTT_neighbor_lemma}
For unary transcripts at $\delta = 1$, If a bead has no free hand, then at least $\alpha + 2$ of its neighbors have to be occpied.
\end{lemma}

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be an oritatami system at $\delta = 1, \alpha =2$ 
Assume $\Xi$ stabilizes transcript until $w[i-1]$. If $w[i]$ is stabilized at an entrance point of tunnel $A$ or $B$, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

\begin{lemma}
\label{TTT_exit}
Let $w[i]$ be a bead which is stabilized in exit of a tunnel.
At $\delta = 1, \alpha =2$, If we assume $S[h..i+1] = bt^{(i-h)}b \quad (h<i)$, then $\#bc(C_{h-1}) \geq \#bc(C_i)$ and $\#bc(C_{i-2}) \geq \#bc(C_i)$.
%if we assume $\#bc(C_{i-2}) - \#bc(C_{i-1}) = a$, then $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$.
On the other hand, if we assume $S[k] = t (k \leq i)$ at $\delta =1, \alpha \geq 3$, then $\#bc(C_{k-1}) > \#bc(C_{k})$.
\end{lemma}

\begin{lemma}
\label{TTT_tunnelC_lemma}
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 2$. We assume $S[h..i+1] = bt^{(i-h)}b \quad (1<h<i)$. If at least one of $w[h+1..i]$ is stabilized by tunnel $C$, then $\#bc(C_{h-3}) > \#bc(C_i)$.
\end{lemma}

Let us first consider cases of $\delta \geq 3, \alpha = 1$. These cases are clearly true because we use lemma\ref{TTT_exit}.

Next, we consider the case of $\delta = 2, \alpha = 1$. We assume there is index $h$ such that $S[h-1..h+1] = bbt$ or $S[h-1..h+1] = tbt$. According to lemma\ref{TTT_entrance_Tab}, if $w[h+1]$ is stabilized by tunnel $A$ or $B$, then $\#bc(C_{h-1}) > \#bc(C_{h})$. Also,  According to lemma\ref{TTT_exit}, if $w[h+1]$ is stabilized by tunnel $C$, then $\#bc(C_{h-1}) > \#bc(C_{h})$, too.
On the other hand, if $S[k..l] = bt^{l-k}b$, then $\#bc(C_{k-1}) \geq \#bc(C_{l})$. Therefore, if there are indices $i$ and $j$ such that $S[i..j+1] = bbt^{(j-i-1)}b$ or $S[i..j+1] = bt^mbt^nb \quad (m + n = j-i-1)$, then $\#bc(C_{i-1}) > \#bc(C_{j})$.

\qed

\begin{figure}
  \begin{center}
    \begin{tabular}{ccc}

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}
          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (0:1) {$p^\prime$};

            \foreach \theta in {0,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,120,-120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel A};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (-60:1) {$p^\prime$};

            \foreach \theta in {-60,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {0,60,120,-120}{
 	            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel B};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (-120:1) {$p^\prime$};

            \foreach \theta in {-120,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {0,60,-60,120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel C};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{All possible tunnel sections: straight, obtuse turn, and acute turn}
    \label{TTT_tunnel}
  \end{center}
\end{figure}

%\begin{figure}
%  \begin{center}
%    \begin{tikzpicture}
%      \draw[thick, gray]
%      (0.5,1)--(0,1)--(0,4)--(4,4)--(4,1)--(3.5,1);
%      \node at(2,1) {Outside};
%      \node at (2,3.2) [rectangle, draw, rounded corners] (enter) {Entrance of a tunnel};
%      \node at(2,1.8) [rectangle, draw, rounded corners] (outside) {Otherwise};
%
%      
%      
%      \draw[thick, gray]
%      (7.5,1)--(7,1)--(7,4)--(11,4)--(11,1)--(10.5,1);
%      \node at(9,1) {Inside of a tunnel};
%      \node at(9,3.2) [rectangle, draw, rounded corners] (next-tunnel) {Successor in inside};
%      \node at(9,1.8) [rectangle, draw, rounded corners] (next-outside) {Successor in outside};
%
%      \draw[->]
%      (outside) edge [bend left, blue] node[left] {$s=b,\Delta=-1$} (enter)
%                   edge [loop left, blue] node[below left] {$s=b,\Delta=\pm 0$} (outside)
%      (enter)   edge [bend left, blue] node[above] {$s=b,\Delta=-1$} (next-tunnel)
%                edge [bend left, blue] node[above] {$s=b,\Delta=-a$} (next-outside)
%                edge [bend left, blue] node[right] {$s=b,\Delta=\pm 0$}  (outside)
%      (next-tunnel) edge[teal]         node[right] {$s=t,\Delta=-a$} (next-outside)
%                edge [loop right, teal]node[right] {$s=t,\Delta=\pm 0$}    (next-tunnel)
%      (next-outside) edge[bend left, teal] node[above] {$s=t,\Delta=+a$}  (outside)
%                edge [bend left, teal]     node[above] {$s=t,\Delta=+(a-1)$}(enter)
%      ;
%      
%    \end{tikzpicture}
%    \caption{Increment on Tunnel A,B}
%    \label{TTT_position}
%  \end{center}
%\end{figure}

%\begin{figure}
%  \begin{center}
%    \begin{tikzpicture}
%      \node at (0,4)   [rectangle, draw,rounded corners] (outside1) {Outside of tunnels};
%      \node at (3.5,4) [rectangle, draw,rounded corners] (enter1)   {Entrance of a tunnel};
%      \node at (7,4)   [rectangle, draw, rounded corners] (inside1) {Inside of a tunnel C};
%      \node at (11,4)  [rectangle, draw, rounded corners] (exit1)   {Inside of a tunnel A or B};
%      \draw[->] (outside1)--(enter1);
%      \draw[->] (enter1)  --(inside1);
%      \draw[->] (inside1) --(exit1);
%      
%      
%      \node at (0,3)   [rectangle, draw,rounded corners] (outside2) {Inside of a tunnel};
%      \node at (3.5,3) [rectangle, draw,rounded corners] (enter2)   {Entrance of a tunnel};
%      \node at (7,3)   [rectangle, draw, rounded corners] (inside2) {Inside of a tunnel C};
%      \node at (11,3)  [rectangle, draw, rounded corners] (exit2)   {Inside of a tunnel A or B};     
%      \draw[->] (outside2)--(enter2);
%      \draw[->] (enter2)  --(inside2);
%      \draw[->] (inside2) --(exit2);
%
%      \node at (0,1)   [rectangle, draw,rounded corners] (outside3) {Outside of tunnels};
%      \node at (3.5,1) [rectangle, draw,rounded corners] (enter3)   {Entrance of a tunnel};
%      \node at (7,1)   [rectangle, draw, rounded corners] (inside3) {Inside of a tunnel C};
%      \node at (11,1)  [rectangle, draw, rounded corners] (exit3)   {Outside of tunnels};     
%      \draw[->] (outside3)--(enter3);
%      \draw[->] (enter3)  --(inside3);
%      \draw[->] (inside3) --(exit3);
%
%      \node at (0,0)   [rectangle, draw,rounded corners] (outside4) {Inside of atunnel};
%      \node at (3.5,0) [rectangle, draw,rounded corners] (enter4)   {Entrance of a tunnel};
%      \node at (7,0)   [rectangle, draw, rounded corners] (inside4) {Inside of a tunnel C};
%      \node at (11,0)  [rectangle, draw, rounded corners] (exit4)   {Outside of tunnels};     
%      \draw[->] (outside4)--(enter4);
%      \draw[->] (enter4)  --(inside4);
%      \draw[->] (inside4) --(exit4);
%    \end{tikzpicture}
%    \caption{Case of Tunnel C}
%    \label{TTT_tunnelC_graph}
%  \end{center}
%\end{figure}

\begin{proof}[lemma \ref{TTT_neighbor_lemma}]
A transcript bead in has predecessor and successor except the first and last beads. If the bead does not have any free hand, then it uses hands with $\alpha$ neighbors. Thus, lemma \ref{TTT_neighbor_lemma} is clearly true.
\end{proof}

\begin{proof}[lemma \ref{TTT_entrance_Tab}]
Fig.\ref{TTT_tunnel_direction} exhibits all the three kinds of entrance of tunnel A, B.
Let $w[i]$ be stabilized at an entrance point of Tunnel $A$ or $B$.
All cases are $\#bc(C_{i-1}) > \#bc(C_{i})$ as follows.

\begin{itemize}
\item{Case of $t_0$}\\
  Let us consider points $n_3,n_4$. At least one of the points $n_3$ or $n_4$ is free because if both of them are occupied, $p^\prime$ is inside of tunnel. If $n_3$ is free, then $p^\prime$ has to be bound to a bead other than $n_1$ due to deterministically stabilize. In this situation, at least three neighbors of $n_1$ are free, that is, $n_1$ has at least one free hand from lemma \ref{TTT_neighbor_lemma}. Hence, $p^\prime$ must be bound to $n_1$. Thus, a case of $t_0$ consumes two hands and it does not supply any binding capabilities.

\item{Case of $t_{\pm 60}$}\\
  In this case, a point $n_4$ or $n_5$ is free, too. If $n_5$ is free, $p^\prime$ has to be bound to $n_1$ or $n_2$. If $n_5$ is occupied, then $n_4$ is free. This time, $n_2$ has some free hands so $p^\prime$ has to be bound to $n_2$.

  In this situation, $p^\prime$ is able to supply a binding capabilities which could $p^\prime$ bind a bead into $n_4$ or $n_5$. However, $n_2$ and $n_3$ are part of a contiguous conformation. According to Jordan curve theorem, any successors of $p^\prime$ cannot reach a point $n_4$ or $n_5$ so this capability is inactive. Thus, a case of $t_{\pm 60}$ is$\#bc(C_{i-1}) > \#bc(C_{i})$.

\item{Case of $t_{\pm 120}$}\\
  Binding capabilities that $p^\prime$ supplies are inactive according to Jordan curve theorem on $n_1$ and $n_2$. Moreover, $p^\prime$ has to be bound to one of $n_3, n_4, n_5$ in order to deterministically stabilize.
Thus, a case of $t_{\pm 120}$ is $\#bc(C_{i-1}) > \#bc(C_{i})$.

\end{itemize}

\begin{figure}
  \begin{center}
    \begin{tabular}{ccc}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
	 \draw[->,white] (120:0.8)--(120:0.1);%for adjustment

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$p^\prime$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)},blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {120,-120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (180:0.9)--(180:0.1);
            \draw[->, blue] (0:0.1)--(0:0.9);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};

          \node at (0,-2) {$t_{0}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}

            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$p^\prime$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)},blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {180,-120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (120:0.9)--(120:0.1);
            \draw[->, blue] (0:0.1)--(0:0.9);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)}, blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {180,-120,120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (60:1.1)--(60:0.1);
            \draw[->, blue] (0:0.1)--(0:1.1);

            \node[transform canvas={shift=(0:1)},below] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Direction into a entrance}
    \label{TTT_tunnel_direction}
  \end{center}
\end{figure}

\end{proof}

\begin{proof}[lemma \ref{TTT_exit}]
Fig.\ref{TTT_tunnel_exit} exhibits all the two kinds of exit of tunnel. 
At least one of points $n_1$ or $n_2$ is free because if both of them are occupied, $p^\prime$ is inside of tunnel.

\begin{paragraph}{}
$\delta = 1, \alpha = 2$\\
Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha = 2$. We assume $S[h..i+1] = bt^{(i-h)}b \quad (h<i)$.
If let $a$ be $\#bc(C_{i-2}) - \#bc(C_{i-1}) = a$, then $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$ as follows. Also, if $h - i > 1$ and let $j$ be $h<j<i$, then $\#bc(C_{j-1}) \geq \#bc(C_{j})$ because any neighbors of $w[j]$ are occupied by beads of tunnel so that any $w[i+1..]$ cannot reach neighbors of $w[j]$. Thus, $\#bc(C_{h-1}) \geq \#bc(C_i)$ and $\#bc(C_{i-2}) \geq \#bc(C_i)$.
\end{paragraph}

\begin{itemize}
\item{Case of $n_1$ and $n_2$ being free}\\
  This case can be regarded the same as entrance. See Fig.\ref{TTT_tunnel_exit} (Left). Predecessor $n_5$ has to be bound to $n_4$ and $n_5$ because both of $n_3$ and $n_4$ have binding capabilities. Hence, $a \geq 2$. This time, $\alpha = 2$, that is, this case $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$.

\item{Case of $n_1$ is occupied}\\
  See Fig.\ref{TTT_tunnel_exit} (Right). If $n_1$ is occupied, then $n_2$ is free so that $n_5$ has to be bound $n_4$. Hence,  $a \geq 1$. This case can supply two binding capabilities but $p^\prime$ can bind to only one of $n_0$ or $n_2$ because $n_0$ or $n_2$ will be occupied by the successor of $p^\prime$. Therefore, this case $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$.

\end{itemize}

\begin{paragraph}{}
$\delta = 1, \alpha \geq 3$\\
Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha \geq 3$. We assume $w[i]$ is stabilized at exit of tunnel.
Any cases $\#bc(C_{i-1}) > \#bc(C_{i})$. Moreover, if $S[k] = t (k \leq i)$, then $\#bc(C_{k-1}) > \#bc(C_{k})$ because both of sides of the path $p$ in Fig.\ref{TTT_inside_tunnel} ($n_1, n_2$) have two free points and one of $n_1, n_2$ is not the predecessor so that it has hand and moreover that $w[k]$ supplies any binding capabilities because its neighbors are occupied by beads of tunnel.
\end{paragraph}

\begin{itemize}
\item{Case of $n_1$ and $n_2$ are free}\\
  In $\alpha \geq 3$, if three neighbors of a bead leave, then it can supply two binding capabilities. Therefore Predecessor $n_5$ has to be bound $n_3$ and $n_4$, and $p^\prime$, too. In this case, at least consumes four bindings and supplies at most two bindings. Thus, it consumes some binding capabilities, totally.

\item{Case of $c$ is occupied}\\
  In this case, $n_4$ leave at least two bindings and $n_3$, $n_1$ also leave at least one binding. Therefore $n_5$ has to be bound $n_3$ and $n_4$, and $p^\prime$ also has to be bound $n_1$ and $n_4$. In this case, at least consumes four bindings and supplies at most one binding. Thus, it consumes some binding capabilities, totally.
\end{itemize}

\begin{figure}
  \centering
    \begin{tabular}{cc}
    \begin{minipage}{0.66\hsize}
    \centering
      \begin{tabular}{cc}
      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
	\draw[white] (0:0) -- (0:1);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

            \foreach \theta in {120,-120,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {0}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }
            \draw (-60:1) circle [radius=0.05];
            \draw (60:1) circle [radius=0.05];

            \draw[->, blue] (180:0.9)--(180:0.1);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {Point $n_1$ and $n_2$ are free};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (0:1);
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-90:2);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$p^\prime$};

            \foreach \theta in {120,-120,180,60}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {0}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }
            \draw (-60:1) circle [radius=0.05];

            \draw[->, blue] (180:0.9)--(180:0.1);

            \node[transform canvas={shift=(0:1)},right] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {Point $n_1$ is occupied};
        \end{tikzpicture}
      \end{minipage}
    \end{tabular}
    \caption{Exit of Tunnel}
    \label{TTT_tunnel_exit}
    \end{minipage}

    \begin{minipage}{0.33\hsize}
    \centering
     \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
            \fill[blue](0,0) circle [radius=0.05];
            \draw[blue] (0:1) circle [radius = 0.05];
            \draw[->] (0:0.1) -- node[above]{$p$} (0:0.9);

            \foreach \theta in {60, -60, 120,-120,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
                \end{tikzpicture}

    \caption{Inside tunnel}
    \label{TTT_inside_tunnel}
    \end{minipage}

    \end{tabular}
\end{figure}
\end{proof}

\begin{proof}[lemma \ref{TTT_tunnelC_lemma}]
Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha = 2$.
Assume $S[h..i+1] = bt^{i-h}b (h<i)$. If at least one of $w[h+1..i]$ are stabilized by tunnel $C$, then only $w[h+1]$ can use tunnel $C$ because if $w[g]$ which is one of $w[h+2..i]$ is stabilized by tunnel $C$, $C_g$ is a terminal.

Let us consider kinds of stabilization $S[h-1..h+1] = tbt$ or $S[h-1..h+1] = bbt$ into follows. In result, $\#bc(C_{h-3}) > \#bc(C_{h+1})$. In addition according to lemma\ref{TTT_exit} $\#bc(C_{h+1}) \geq \#bc(C_{i})$. Thus, $\#bc(C_{h-3}) > \#bc(C_{i})$.
\\

\subsubsection{Case of $S[h-1..h+1] = tbt$}
Fig.\ref{TTT_tunnelC_enter_usingTunnel} exhibits all the two kinds of stabilization depending on structures of tunnel C.

\begin{itemize}
\item{Left of Fig.\ref{TTT_tunnelC_enter_usingTunnel}}\\
  In this figure, Bead $n_4$ has at least one binding so that $w[h-1]$ has to bound $n_4$. Moreover, $w[h]$ has to bound one of $n_1, n_2, n_3$ in order to stabilize deterministically. On the other hand, $w[h+1]$ can supply two bindings but free neighbors of $w[h+1]$ are two points. One of them is occupied by a successor. Therefore $w[h+1]$ can only bind one of $n_5, n_6$, that is, $w[h+1]$ supplies at most one binding. Thus, this case $\#bc(C_{h-1}) > \#bc(C_{h+1})$.

\item{Right of Fig.\ref{TTT_tunnelC_enter_usingTunnel}}\\
  This cases are divided on number of capabilities that $w[h-1]$ consumes.
  \begin{itemize}
  \item[-]{$w[i]$ does not consume any bindings}\\
  According to lemma\ref{TTT_exit}, $\#bc(C_{h-3}) \geq \#bc(C_{h-1})$ because of $S[h-1] = t$.
    $w[h]$ has to bound one of $n_1, n_2, n_3$ in order to stabilize deterministically so that  $\#bc(C_{h-1}) > \#bc(C_{h})$.
    $w[h+1]$ has to be bound to $w[h-1]$ because $w[h-1]$ has bindings, that is, $w[h+1]$ consumes at least one hand and supplies at most one hand so that $\#bc(C_{h}) \geq \#bc(C_{h+1})$. Thus, in this cases $\#bc(C_{h-3}) > \#bc(C_{h+1})$.
%     This time, let us consider either $n_4$ is occupied or not. If $n_4$ is occupied, then $w[h-1]$ has no active bindings that is this situation consumes some binding capabilities. If $n_4$ is free and $w[h+2]$ is stabilized in $n_4$, then $w[h-1]$ has to bind $w[h+2]$. Therefore, In this case, stabilization of $w[h-1..h+2]$ consumes some bindings. If $n_4$ is free and $w[h+2]$ is stabilized except $n_4$, then this oritatami system has to use two binding capabilities in order to bind $w[h+2]$. Therefore, in this case consumes some bindings. Thus in this cases $\#bc(C_{h-1}) > \#bc(C_{h+2})$.

  \item[-]{$w_{h+1}$ consumes one binding}\\
    In this case, $w_{h}$ has to be bound one of $n_1, n_2, n_3$. In addition, $w[h-1]$ and $w[h+1]$ are not supply any bindings. Thus, in this cases consume some binding capabilities.
  \item[-]{$w_i$ consumes two bindings}\\
    In this case, $w[h-1]$ already consumes two binding. $w[h]$ has to be bound. $w[h+1]$ supplies two bindings. Thus, in this cases $\#bc(C_{h-1}) > \#bc(C_{h+1})$.

  \end{itemize}
\end{itemize}
\begin{figure}
  \centering
    \begin{tabular}{cc}

      \begin{minipage}{0.48\hsize}
      \centering
        \begin{tikzpicture}

          \fill[transform canvas={shift=(0:0)}](0,0) circle [radius=0.1];

          \foreach \theta in {60,-60,120,-120,180}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

          \fill[blue](0:1) circle [radius=0.05];
          \draw[->, blue] (0:0.1)--(0:0.9);

          \node[below] at (-60:1) {$n_4$};

          \begin{scope}[shift=(0:2)]
            \fill[blue](0,0) circle [radius=0.05];

            \foreach \theta in {120,60,-60,0}{
             \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \draw[->, blue] (180:0.9)--(180:0.1);
            \draw[->, blue] (-120:0.1)--(-120:0.9);

            \node[below] at (180:1) {$w[h-1]$};
            \node[above] at (0:0) {$w[h]$};
            \node[right] at (-120:1) {$w[h+1]$};

            \node[above] at (120:1) {$n_3$};
            \node[above] at (60:1) {$n_2$};
            \node[right] at (0:1) {$n_1$};

            \begin{scope}[shift=(-120:1)]
              \fill[blue](0,0) circle [radius=0.05];
              \foreach \theta in {-120,-60}{
                \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
              }

              \node[right] at (-120:1) {$n_5$};
              \node[right] at (-60:1) {$n_6$};
            \end{scope}

          \end{scope}

          \node at (1.2,-2.5) {Pattern 1};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.48\hsize}
      \centering
        \begin{tikzpicture}

          \fill(0,0) circle [radius=0.1];

          \foreach \theta in {60,0,120,-120,180}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

          \draw[->, blue] (-60:0.1)--(-60:0.9);

          \begin{scope}[shift=(-60:1),shift=(0:1)]
            \fill[blue](0,0) circle [radius=0.05];
            \fill[blue](180:1) circle [radius=0.05];

            \foreach \theta in {120,60,0,-60}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \draw[->, blue] (180:0.9)--(180:0.1);
            \draw[->, blue] (-120:0.1)--(-120:0.9);

            \node[below] at (180:1) {$w[h-1]$};
            \node[above] at (0:0) {$w[h]$};
            \node[right] at (-120:1) {$w[h+1]$};

            \node[above right] at (120:1) {$n_3$};
            \node[right] at (60:1) {$n_2$};
            \node[right] at (0:1) {$n_1$};

            \begin{scope}[shift=(-120:1)]
              \fill(0,0) circle [radius=0.05];
              \foreach \theta in {180,-120,-60}{
                \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
              }

              \node[right] at (180:1) {$n_4$};
              \node[right] at (-120:1) {$n_5$};
              \node[right] at (-60:1) {$n_6$};
            \end{scope}

          \end{scope}

          \node at (1.25,-4) {Pattern 2};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Case of $S[h-1..h+1] = tbt$}
    \label{TTT_tunnelC_enter_usingTunnel}
\end{figure}

\subsubsection{Case of $S[h-1..h+1] = bbt$}
Let us consider number of consumed by $w[h-1]$ (Fig.\ref{TTT_tunnelC_enter_usingBond}).

\begin{itemize}
\item{$w[h-1]$ consumes one binding}\\
  In this situation, $w[h-1]$ supplies one active binding whereas $w[h+1]$ consumes this binding. In addition, $w[h]$ has to bound to one of $n_1, n_2, n_3$.
  Thus, in this cases consume some binding capabilities.

\item{$w[h-1]$ consumes two bindings}\\
  In this case, $w[h-1]$ already consumes two binding. $w[h]$ has to be bound. $w[h+1]$ supplies at most two bindings. Thus, in this cases consume some binding capabilities.
\end{itemize}

\begin{figure}
  \begin{center}
        \begin{tikzpicture}
        \draw[->,white] (-70:0.1) -- (-70:3);
          \foreach \theta in {60,-60,120,0}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

          \fill[blue](180:1) circle [radius=0.05];
          \fill[blue](0:0) circle [radius=0.05];

          \draw[transform canvas={shift=(180:1)}, blue] (105:1) edge[bend right] (105:0);
          \draw[->, blue] (180:0.9) -- (180:0.1);
          \draw[->, blue] (-120:0.1) -- (-120:0.9);

          \node[below left] at (180:1) {$w[h-1]$};
          \node[above] at (0:0) {$w[h]$};
          \node[below] at (-120:1) {$w[h+1]$};

          \node[right] at (120:1) {$n_3$};
          \node[right] at (60:1) {$n_2$};
          \node[right] at (0:1) {$n_1$};

          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];
            \foreach \theta in {-120,-60,180}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[h-1..h+1] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/////////////////////////////////////////////////
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{On structures provided by a unary and $\delta = 1$ oritatami system}
\begin{theorem}[$\delta= 1, \alpha=2$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 2$. It can yield infinite structures but they are only zig-zag conformation.
\end{theorem}

\begin{proof}
By Tunnel Troll Theorem, any tunnel sections which represented in $bbt^+$ or $bt^+bt^+$ consume binding capabilities. If the sequence $S$ is free from any subsequence of the form $bt^+bt^+$, then it can factorize as $S = u_1 u_2 u_3 \cdots$ for some $u_1 , u_2 , u_3 , \cdots \in \{b\} \cup bbt^+$. Assume the length of $\sigma$ is $n$, seed supplies at most $2n$ binding capabilities. Therefore formula \ref{TTT_only_bond} hold.

\begin{eqnarray}
  \exists i \in \mathbb{N} \quad s.t. \quad u_{i-1} , u_i , u_{i+1} , u_{i+2} , \cdots \in \{ b \}
  \label{TTT_only_bond}
\end{eqnarray}

Let us represent $S$ as $S[i.i+1...] = v_i v_{i+1} v_{i+2} \cdots$ for some $v_i, v_{i+1}, v_{i+2}, \cdots \in \{ a, o\}$ where if $v_k$ is $a$, then $v_{k+1}$ is bound to $v_{k-1}$, if $v_k$ is $o$, then $v_{k+1}$ is NOT bound to $v_{k-1}$.

Let us consider the case of $v_k$ is $o$. See Fig.\ref{TTT_case_of_o}. $w[i-1]$ consumes some binding capabilities because $v_{i-1}$ is $b$. If the number of $w[i-1]$'s bindings is one binding, then $w[i+1]$ has to be bound except $n_1$ or $n_2$ so that $w[i+1]$ must consumes two bindings except the case of $n_1$ and $n_2$ are occupied and $w[i]$ consumes at least one binding. If $n_1$ and $n_2$ are occupied, then $w[i-1]$'s bindings are inactive, that is, $w[i-1]$ consumes two binding capabilities. Therefore, this case consumes binding capabilities. If $w[i-1]$ dose Not have any bindings, then $w[i-1]$ already consumes two bindings. In addition, $w[i]$ and $w[i+1]$ consume at least one binding. Therefore this case consumes binding capabilities. Thus, the formula \ref{TTT_only_a} hold and according to the formula \ref{TTT_only_bond} and the formula \ref{TTT_only_a}, the formula \ref{TTT_structure} is hold. Thus, in this case, oritatami system can yield infinite structures but they are only zig-zag conformation.

\end{proof}

\begin{eqnarray}
  \exists j \in \mathbb{N} \quad s.t. \quad u_j , u_{j+1} , u_{j+2} , \cdots \in \{ a \}
  \label{TTT_only_a}
\end{eqnarray}

\begin{eqnarray}
  | S | > \forall m \in \mathbb{N} \quad \to \quad \exists n \in \mathbb{N} \quad s.t. \quad S[n], S[n+1], \cdots \in \{ a \}
  \label{TTT_structure}
\end{eqnarray}

\begin{figure}
  \begin{center}
    \begin{tikzpicture}

      \fill[transform canvas={shift=(-120:1)}](0,0) circle [radius=0.1];

      \fill[blue](0:0) circle [radius=0.05];
      \fill[blue](0:1) circle [radius=0.05];
      \fill[shift=(0:1), blue](40:1) circle [radius=0.05];

      \draw[->, blue] (-120:1) edge[bend left] (-120:0.1);
      \draw[->, blue] (0:0.1) -- (0:0.9);
      \draw[->,transform canvas={shift=(0:1.2)}, blue] (50:0) -- (47:0.8);

      \draw[shift=(60:1)](-0.05,-0.05) rectangle (0.05,0.05);
      \draw[shift=(-60:1)](-0.05,-0.05) rectangle (0.05,0.05);

      \node[below] at (-120:1) {$w[i-2]$};
      \node[above] at (0:0) {$w[i-1]$};
      \node[below] at (0:1) {$w[i]$};
      \node[right, shift=(0:1)] at (40:1) {$w[i+1]$};

      \node[above] at (60:1) {$n_1$};
      \node[below] at (-60:1) {$n_2$};

    \end{tikzpicture}
    \caption{Case of $S[i]$}
    \label{TTT_case_of_o}
  \end{center}
\end{figure}

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only finite structures whose size is $\mathcal{O}(n)$.
\end{theorem}

\begin{lemma}
\label{TTT_a3_2b_lemma}
 Let $p$ be a point whose neighbors is occupied at least two point. If $w[i]$ is not stabilized and $w[i-1]$ includes neighbors of $p$, then $w[i]$ is stabilized at $p$ with at least one bond, $w[i]$ is stabilized at another point of $p$ otherwise with at least two bond except any neighbors of $p$ is occupied.
\end{lemma}

\begin{proof}[proof of lemma]
Assume the transcript is stabilized until $w[i-1]$. One of neighbors of $p$ is not $w[i-1]$ where this bead regards $n_1$. If $w[i-1]$ include neighbors of $p$ and $w[i]$ is stabilized at another point of $p$ with one bond. Then, any neighbors do not have bond without $w[i-1]$. Neighbors of $n_1$ have to be occupied at least five according to lemma \ref{TTT_neighbor_lemma} and two of them include neighbors of $p$ where each of them regards $n_2, n_3$. In the same way, five neighbors of $n_2$ and $n_3$ are occupied and each of one of them includes neighbors of $p$ where they regard $n_4, n_5$. one of $n_5$'s neighbors includes neighbors of $p$ where it regards $n_6$. Then, any neighbors of $p$ are occupied. That is, if some neighbors of $p$ are free, then there exists a bead which has bonds in neighbors.
\end{proof}

\begin{proof}
Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is, when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
Let us assume $w[i]$ is able to be stabilized with using one hand. Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$. Then, $w[i]$ can be also stabilized at $n_3$.

\begin{paragraph}{Case of straight}
\begin{itemize}
\item[-] Case of $n_3$ is free\\
According to assumption, $w[i]$ uses only one hand. Therefore, any neighbors of $n_3$ are occupied according to the lemma\ref{TTT_a3_2b_lemma}. $n_3$ and the point which is stabilized $w[i]$ are free so that $n_1$ has some bond by lemma\ref{TTT_neighbor_lemma}. Accordingly, this situation is non-deterministic. Thus, $n_3$ and $n_4$ have to be occupied because of symmetry.

\item[-] Otherwise\\
Because of  $S[i] = b$, at least one of $n_1$ and $n_2$ have to be free. Let us regard that $n_1$ is free. Neighbors of $n_1$ have to be occupied and at least two neighbors of $n_{-1}$ have to be free for $n_1$ and $w[i]$. According to lemma\ref{TTT_neighbor_lemma}, $n_{-1}$ have some hand. Therefore $w[i]$ can be also stabilized $n_1$, that is, this situation is non-deterministic. Thus, one of $n_3$ and $n_4$ has to be free.
\end{itemize}

Therefore, this case is false.
\end{paragraph}

\begin{paragraph}{Case of obtuse}
\begin{itemize}
\item[-] Case of $n_3$ is free\\
Any neighbors of $n_3$ have to be occupied but the point which is stabilized $w[i]$ is free. Thus $n_3$ has to be occupied.
\item[-] Case of $n_4$ is free\\
According to lemma\ref{TTT_a3_2b_lemma}, $n_2$ has to be occupied because $n_4$ is free. Also $n_0$ has to be occupied from lemma\ref{TTT_neighbor_lemma}. Thus, only one of $n_0, n_3$ leave some hands or both of them do not leave any hands because $w[i]$ use only one bond.\\
If $n_0$ has some hands, then $n_3$ does not have any hands so that $n_{-3}$ is occupied. Also $n_{-3}$ must not have any hands so that $n_{-2}$ is occupied and also $n_{-1}$ is occupied. Therefore any neighbors of $w[i]$ are occupied so that $w[i+1]$ cannot provide.\\
If $n_3$ has some hands, then $n_0$ does not have any hands so that $n_{-1}$ is occupied. In the same previous way, any $n_{-2}, n_{-3}$ are occupied.  Therefore any neighbors of $w[i]$ are occupied.\\
If both of $n_0, n_3$ do not have any hands, then both of $n_{-1}, n_{-3}$ are occupied. If one of $n_{-1}, n_{-3}$ has some hands, the other does not have any hands so that $n_{-2}$ is occupied. If both of $n_{-1}, n_{-3}$ do not have any hands, $n_{-2}$ has to be occupied and $n_{-2}$ has some hands.Therefore any neighbors of $w[i]$ are occupied so that $w[i+1]$ cannot provide.\\
Thus $n_3$ has to be occupied in order to yield infinite structures.
\item[-] Case of $n_2$ is free\\
Any neighbors of $n_2$ have to be occupied so that $n_0$ is occupied. Any neighbors of $n_0$ except $n_2$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_2$ has to be occupied.
\item[-] Case of $n_0$ is free\\
Any neighbors of $n_0$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
\end{itemize}

Therefore, any situations contradict $S[i] = b$.

\end{paragraph}

\begin{paragraph}{Case of acute}
\begin{itemize}
\item[-] Case of $n_4$ is free\\
$n_4$ and a point which is stabilized $w[i]$ are free so that  $w[i-2]$ has some hands according to lemma\ref{TTT_neighbor_lemma}.
However, $w[i]$ can be also stabilized $n_4$ in this case. Thus, $n_4$ has to be occupied.
\item[-] Case of $n_2$ is free\\
According to lemma\ref{TTT_a3_2b_lemma}, $n_0$ has to be occupied. $n_1$ has to be also occupied because of lemma\ref{TTT_neighbor_lemma}.
We consider this case just like case of obtuse and that $n_4$ is free. Then if $w[i-2]$ binds $w[i]$, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. If $n_1$ binds $w[i]$, this case is same. Also if $n_1$ and $w[i-2]$ do not have any hand, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. Therefore, $w[i+1]$ cannot be provided.
\item[-] Case of $n_0$ is free\\
Any neighbors of $n_0$ have to be occupied so that $n_{1}$ is occupied. Any neighbors of $n_{1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
\item[-] Case of $n_1$ is free\\
Any neighbors of $n_1$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_1$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_1$ has to be occupied. 
\end{itemize}

Therefore, any situations contradict $S[i] = b$.
\end{paragraph}\\

Hence, assumption that w[i] is able to be stabilized with using one hand is false.
Therefore, when w[i] is stabilized, w[i] uses at least two hands.
\end{proof}

\begin{figure}
  \begin{center}
  \begin{tabular}{c c c}
 \begin{minipage}{0.3\hsize}
 \centering
    \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](0:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (0:0.1) -- (0:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (0:1) {$w_{[i]}$};

	\foreach \theta in {60,-60,120,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	 \draw [shift=(60:1)] (0.95,-0.05) rectangle (1.05,0.05);
 	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(0:1)] at (60:1) {$n_{-1}$};
    \end{tikzpicture}
    \end{minipage}

    \begin{minipage}{0.3\hsize}
    \centering
     \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](60:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (60:0.1) -- (60:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (60:1) {$w_{[i]}$};

	\foreach \theta in {0,-60,120,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	\draw [shift=(60:1)] (0.95,-0.05) rectangle (1.05,0.05);
	\draw [shift=(60:1), shift=(60:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(60:1), shift=(120:1)] (-0.05,-0.05) rectangle (0.05,0.05);

 	\node[right] at (0:1) {$n_0$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(0:1)] at (60:1) {$n_{-1}$};
	\node[above, shift=(60:1)] at (60:1) {$n_{-2}$};
	\node[above, shift=(60:1)] at (120:1) {$n_{-3}$};
    \end{tikzpicture}
    \end{minipage}

    \begin{minipage}{0.3\hsize}
        \centering
     \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](120:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (120:0.1) -- (120:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (120:1) {$w_{[i]}$};

	\foreach \theta in {0,60,-60,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	\draw [shift=(120:1), shift=(60:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(120:1), shift=(120:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(120:1), shift=(180:1)] (-0.05,-0.05) rectangle (0.05,0.05);

 	\node[right] at (0:1) {$n_0$};
	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(120:1)] at (60:1) {$n_{-1}$};
	\node[above, shift=(120:1)] at (120:1) {$n_{-2}$};
	\node[above, shift=(120:1)] at (180:1) {$n_{-3}$};
    \end{tikzpicture}
    \end{minipage}
    \end{tabular}
    \caption{All possible directions of $w[i]$: straight, obtuse, acute.}
    \label{TTT_a3_w}
  \end{center}
\end{figure}

\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only finite structures whose size is $\mathcal{O}(n^2)$.
\end{theorem}

\begin{lemma}
\label{TTT_a4_2b_lemma}
Any beads which are already stabilized by some bonds use at least two bonds.
\end{lemma}

\begin{proof}[proof of lemma]
Let us consider when $w[i]$ is stabilized by only one bond. See Fig.\ref{TTT_a4_lemma_fig}. According to lemma\ref{TTT_neighbor_lemma}, if $n_3$ is free, $w[i-2]$ has some hands. Thus, $n_4$ has to be occupied in order to stabilize deterministically. Moreover, also $n_2$ has to be occupied for deterministic and also $n_0, n_1$. $n_1$ has some hands because $n_3$ is free. Therefore, $w[i]$ is stabilized at $n_3$ and it has to use at least two hands. It contradict assumption.
\end{proof}

\begin{proof}
According to lemma\ref{TTT_a4_2b_lemma}, when $w[i]$ is stabilized, it has to use at least two bonds. Let us consider when a bead $w[i]$ which is the first bead out of $\hexagon_{w[-n+1]}^n$ is stabilized. See Fig.\ref{TTT_a4_first}. any $n_0, n_1, n_3, n_5$ is free because if some of them is occupied, $w[i]$ is not the first bead out of $\hexagon_{w[-n+1]}^n$. At least two neighbors of $w[i]$ except predecessor have to be occupied in order to bind. In this case, a point which is able to put a bead is only $n_2$. Therefore, any transcript cannot be stabilized in out of $\hexagon_{w[-n+1]}^n$. Hence oritatami system can yield only a finite structure whose size is $\mathcal{O}(n^2)$ in $\delta =1, \alpha = 4$.
%lemma???a=4??#bc(C_i) >= #bc(C_{i+1})? ?upper?left
\end{proof}

\begin{figure}
 \centering
    \begin{tikzpicture}
	\fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \draw[->] (180:0.9) -- (180:0.1);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (0:1) {$n_0$};

	\foreach \theta in {0,60,-60,120,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
 	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[below] at (-120:1) {$n_4$};
    \end{tikzpicture}
    \caption{$\alpha = 4$: when $w[i]$ is stabilized}
    \label{TTT_a4_lemma_fig}
\end{figure}

\begin{figure}
 \centering
    \begin{tikzpicture}
	\fill[shift=(-120:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0), blue] (0,0) circle [radius=0.05];

      \draw[dashed] (180:2) -- (0:2);
      \draw[dashed, shift=(-60:1)] (180:2.5) -- (0:1.5);
      \draw[->, blue] (-120:0.9) -- (-120:0.1);

	\node[below] at (-120:1) {$w_{[i-1]}$};
	\node[above] at (180:0) {$w_{[i]}$};
	\node[above] at (0:1) {$n_0$};

	\foreach \theta in {0,60,-60,120,180}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}

 	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[above] at (180:1) {$n_5$};

	\node at (0:2.5) {$n+1$};
	\node[shift=(-60:1)] at (0:2) {$n$};
    \end{tikzpicture}
    \caption{the first bead out of $\hexagon_{w[-n+1]}^n$}
    \label{TTT_a4_first}
\end{figure}

%end of d = 1
%//////////////////////////////////////////////////////% l.177 - 1312
%//////////////////////////////////////////////////////%

\section{Case of $\delta \geq 2$}

\subsection{$\alpha = 1$, unary}

Let the point where the first transcript bead was fixed be $p$ and let $n=|\mathrm{seed}|+1$. We will argue about the situation when the first bead is stabilized outside $\hexagon_p^n$ (a hexagon of radius $n$). Let this be the $i$th bead of the transcript. Without loss of generality, we can translate the origin $(0,0)$ to the coordinates of bead $i-1$ (which is still in $\hexagon_p^n$), and we can assume that the bead outside the hexagon is fixed at $(1,1)$ (see Fig.~\ref{fig:hexagonOut}).
\begin{figure}
	\centering
	\includegraphics[width=0.3\linewidth]{./Fig/hexagonOut}
	\caption{$N_p^n$ and the position $(1,1)$ of the first bead fixed outside of it.}
	\label{fig:hexagonOut}
\end{figure}

In the elongation that places bead $i$ at $(1,1)$ there are two possibilities:
\begin{itemize}
	\item $i$ forms a bond with a bead at $(1,0)$.
	\item  $i$ does not bond to anything and $i+1$ is at $(2,1)$ bonding with a bead at $(2,0)$. If there is no bead at $(1,0)$, then placing $i$ at $(1,0)$ instead of $(1,1)$ results in the same number of bonds, leading to nondeterminism. Therefore, there has to be a bead at $(1,0)$ and it is inert, otherwise it would bond to $i$. This is analogous to case 1. below.%as in Fig.~\ref{fig:hexagonOut1}.
	\begin{figure}
		\centering
		\includegraphics[width=0.2\linewidth]{./Fig/hexagonOut1}
		%\caption{}
		\label{fig:hexagonOut1}
	\end{figure}

\end{itemize}

%The only position with which $(1,1)$ can form a bond is $(1,0)$. This means that there is a bead at $(1,0)$, which bonds to bead $i$, otherwise there are other conformations in which beads $i$ and $i+1$ add one bond to the conformation, making the behavior nondeterministic.

The next bead, $i+1$, can be fixed at $(2,1)$ or at $(0,1)$ as all other possibilities result in nondeterministic behavior immediately, so we have two cases.

\begin{enumerate}
	\item bead $i+1$ is fixed at $(2,1)$ and can bond with a bead at $(2,0)$. Now consider bead $i+2$. For $i+1$ to be fixed at $(2,1)$, $i+2$ needs to form a bond somewhere, otherwise $i+2$ could go to $(2,1)$ forming the bond with the bead at $(2,0)$ and there would be two conformations with the maximal $1$ bond. The only possibility is that there is a bead at $(3,0)$ and $i+2$ can bond with it when placed at $(3,1)$. We can apply the same argument inductively: there is some $m\geq 0$ such that grid points $(\ell,0)$ are occupied by active beads, for all $\ell\in \{2,\dots,2+m\}$, and there is no bead at $(3+m,0)$. Such an $m$ exists, and it is not greater than $n$. Then, bead $i+\ell$ is fixed at $(\ell+1,1)$ and bonds with $(\ell+1,0)$. However, bead $i+2+m$ cannot be fixed anywhere, because $i+2+m$ and $i+3+m$ can only add one bond to the conformation, and that is possible either with $i+2+m \rightarrow (2+m,1)$, $i+3+m \rightarrow (3+m,1)$ or with $i+2+m \rightarrow (2+m,2)$, $i+3+m \rightarrow (2+m,1)$. 
	\begin{figure}
		\centering
		\includegraphics[width=0.6\linewidth]{./Fig/hexagonOut2}
		\caption{}
		\label{fig:hexagonout2}
	\end{figure}

	\item bead $i+1$ is fixed at $(0,1)$. This is only possible if
	\begin{enumerate}
		\item there is an inactive bead at $(-1,0)$ and an active one at $(-2,0)$. This case is symmetrical to (1).
		\item there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ has two bonds, just as the original conformation.
		\item there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

	\end{enumerate}
\end{enumerate}

\begin{figure}
	\centering
	%\includegraphics[width=0.3\linewidth]{./hexagonOut1}
	%\hspace{10mm} %
	\includegraphics[width=0.6\linewidth]{./Fig/hexagonOut3}

	\caption{}
	\label{fig:hexagonOut2}
\end{figure}

\subsection{In the case of $\delta = 3$}

We discuss when the length of seed is $n$. If the length of seed is $n$, the seed will be in an equilateral hexagon with a radius of $n$ when it is given. We will begin our disscussion by considering the moment when a first bead is fixed on a side of an equilateral hexagon with a radius of $n+1$. Let us assume that a fixed bead on a side of an equilateral hexagon with a radius of $n$ is able to make a hydrogen bond. If it is able to make only two hydrogen bonds, it will be nondeterministic because there are two possibilities that it makes two structures in Fig.5. Therefore, it needs to make three hydrogen bonds to determine the only structure.\\

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
        \draw (60 : 1.0) -- (120 : 1.0);
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \foreach \theta in {60}{
        \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
        \draw (60 : 1.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \foreach \theta in {60}{
        \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted](0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
  \end{center}
\end{figure}

There are the two cases in which a first bead is fixed on a side of an equilateral hexagon with a radius of $n+1$ and it makes three hydrogen bonds in Fig.6. If it makes three hydrogen bonds once such as Fig6., it will make three hyrogen bonds forever to be deterministic. This becomes nondeterministic finally, such as in Fig.7., when it reaches a corner of an equilateral hexagon. Thus, the fixed bead on a side of an equilateral hexagon with a radius of $n$ has to make a hydrogen bond with a bead in an equilateral hexagon.

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (3.5, 0.0);
      \fill (3.5, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (3.5, 0.866);
      \fill (3.5, 0.866) node [right] {$n+1$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \fill (0,0) circle [radius = 0.05];
        \foreach \theta in {0, 60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
          \draw (0 : 0.0) -- (0 : 2.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 3.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-1.5, 0.0) -- (2.5, 0.0);
      \fill (2.5, 0.0) node [right] {$n$};
      \draw [dashed] (-1.5, 0.866) -- (2.5, 0.866);
      \fill (2.5, 0.866) node [right] {$n+1$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {0, 60, 120, 180}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \fill (0,0) circle [radius = 0.05];
        \foreach \theta in {120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
          \draw (0 : 0.0) -- (0 : 2.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (180 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (2.0, 0.0);
      \fill (3.0, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (2.5, 0.866);
      \fill (3.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-0.5, 1.732) -- (3.0, 1.732);
      \fill (3.0, 1.732) node [right] {$n+2$};
      \fill (60 : 1.0) circle [radius = 0.05];
      \draw (-0.1,-0.1) -- (0.1,0.1);
      \draw (0.1,-0.1) -- (-0.1,0.1);
      \begin{scope}[shift = (120 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (300 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw (0.0, 0.0) -- (2.0, 0.0);
        \fill (2.0, 0.0) circle [radius = 0.05];
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dashed] (0 : 0.0) -- (300 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
      \fill (0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \fill (0.0, 0.0) circle [radius = 0.05];
          \draw (0 : 0.0) -- (240 : 1.0);
          \draw (0 : 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (2.0, 0.0);
      \fill (3.0, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (2.5, 0.866);
      \fill (3.0, 0.866) node [right] {$n+1$};
      \draw (-0.1,-0.1) -- (0.1,0.1);
      \draw (0.1,-0.1) -- (-0.1,0.1);
      \fill (60 : 1.0) circle [radius = 0.05];
      \begin{scope}[shift = (120 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (300 : 1.0);
        \draw (0.0, 0.0) -- (3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dashed] (0 : 0.0) -- (300 : 1.0);
        \fill (60 : 1.0) circle [radius = 0.05];
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
  \end{center}
\end{figure}

Then, we will discuss the moment after the first the is fixed on a side of an equilateral hexagon with a radius of $n+1$. Let us assume that the bead is able to make a hydrogen bond. If it is able to make only two hydrogen bonds, it will be nondeterministic because there are possibilities that it makes two structures in Fig.8. Hence, it needs to make three hydrogen bonds to determine the only structure. It needs to be a right or left figure in Fig.9. to make three hydrogen bonds.\\

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
        \begin{scope}[shift=(0 : 0.0)]
          \draw (-0.1,-0.1) -- (0.1,0.1);
          \draw (0.1,-0.1) -- (-0.1,0.1);
          \draw (0 : 0.0) -- (60 : 1.0);
        \end{scope}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \draw [dashed] (-1.0, 2.598) -- (2.0, 2.598);
      \fill (2.0, 2.598) node [right] {$n+3$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 2.0)]
        \fill(0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {120, 180, 240}{
          \fill (0.0, 0.0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (180 : 1.0);
        \draw (0 : 0.0) -- (240 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \draw (0.0, 0.0) -- (60 : 1.0);
          \draw [dotted] (0.0, 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
        \begin{scope}[shift=(0 : 0.0)]
          \draw (-0.1,-0.1) -- (0.1,0.1);
          \draw (0.1,-0.1) -- (-0.1,0.1);
          \draw (0 : 0.0) -- (60 : 1.0);
        \end{scope}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \draw [dashed] (-1.0, 2.598) -- (2.0, 2.598);
      \fill (2.0, 2.598) node [right] {$n+3$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 2.0)]
        \fill(0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {120, 180, 240}{
          \fill (0.0, 0.0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (180 : 1.0);
        \draw [dotted] (0 : 0.0) -- (240 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \draw [dotted] (0.0, 0.0) -- (60 : 1.0);
          \draw (0.0, 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
  \end{center}
\end{figure}

 If it makes three hydrogen bonds once such as Fig.9., it will make three hyrogen bonds forever to be deterministic. This becomes nondeterministic finally, such as in Fig.7., when it reaches a corner of an equilateral hexagon. Therefore, the first fixed bead on a side of an equilateral hexagon with a radius of $n+1$ has to make a hydrogen bond with a molecule on a side of an equilateral hexagon with a radius of $n$.

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \draw [dashed] (-0.5, 0.0) -- (4.5, 0.0);
      \fill (4.5, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (4.5, 0.866);
      \fill (4.5, 0.866) node [right] {$n+1$};
      \fill (2.0, 0.0) circle [radius = 0.05];
      \fill (3.0, 0.0) circle [radius = 0.05];
      \fill (4.0, 0.0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \fill (1.0, 0.0) circle [radius = 0.05];
        \fill (2.0, 0.0) circle [radius = 0.05];
        \fill (3.0, 0.0) circle [radius = 0.05];
        \draw (0.0, 0.0) -- (3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 3.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 4.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \draw [dashed] (-3.5, 0.0) -- (1.0, 0.0);
      \fill (1.0, 0.0) node [right] {$n$};
      \draw [dashed] (-3.5, 0.866) -- (1.0, 0.866);
      \fill (1.0, 0.866) node [right] {$n+1$};
      \fill (-1.0, 0.0) circle [radius = 0.05];
      \fill (-2.0, 0.0) circle [radius = 0.05];
      \fill (-3.0, 0.0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \fill (-1.0, 0.0) circle [radius = 0.05];
        \fill (-2.0, 0.0) circle [radius = 0.05];
        \fill (-3.0, 0.0) circle [radius = 0.05];
        \draw (0.0, 0.0) -- (-3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : -1.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : -2.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : -3.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
  \end{center}
\end{figure}

Accordingly, it is Fig.10. when a first bead is fixed on a side of an equilateral hexagon with a radius of $n+1$.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift=(60 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift=(0 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \caption{}
  \end{center}
\end{figure}

It has to be a right or left figure in Fig.6. when the first bead fixed on a side of an equilateral hexagon with a radius of $n+1$ make a hydrogen bond with a bead on a side of an equilateral hexagon with a radius of $n$. Fig.6. becomes nondeterministic finally, such as in Fig.7., when it reaches a corner of an equilateral hexagon. It is clear that all beads on a side of an equilateral hexagon with a radius of $n+1$ has to make a hydrogen bond with a bead on a side of an equilateral hexagon with a radius of $n$.  It is impossible to fix molecules on a side of an equilateral hexagon with a radius of $n+2$ and determine the only structure when there are only beads which are able to make a hydrogen bond in an equilateral hexagon with a radius of $n$.

\bibliographystyle{splncs03}
\bibliography{tamc2019.bib}

\end{document}