### 2018-12-02 arity1.tex
Rewrote the Chichen Itza part
The previous version was hard to follow, because one got lost in the details. I think now it is much more readable and looks more organized.

The only modification to the main file, paper.tex are some package declarations.

\subsection{$\alpha = 1$, arbitrary alphabets}
First we present a lower bound construction for arity $\alpha=1$ systems. At $\alpha=1$, allowing delay $\delta=2$ increases the power of OS compared with $\delta=1$, if at least four bead types are allowed. We demonstrate this with an infinite family of OS, which fold deterministically a transcript of length $\frac{(n-1)^2}{4}$ starting from a given seed of length $n$.

Consider the following $\delta=2$, $\alpha=1$ system with bead types $\{0,1,2,3,4\}$ and attraction rules $\{(1,1),(2,2), (3,3),(4,4)\}$. Let the seed $\sigma$ be a conformation of a $4k+1$ long bead sequence of the form $(1020)^k0$, such that bead $\sigma[i]$ of the seed is stabilized at point $(i,0)$, for all $1\leq i\leq 4k-1$. Bead $4k$ is at $(4k-1,-1)$ and bead $4k+1$ is at $(4k,0)$.

\noindent The transcript is
\begin{center}

	%\scriptsize
	\begin{tabular}{llll}
		& $k$ odd &  & $k$ even\\
		$w=$\hspace{1cm} & $(2413)^{k-1}241$ & \hspace{1cm} &  $(2413)^{k-1}241$\\
		     & $(4231)^{k-1}4$   & &  $(4231)^{k-1}4$  \\
		     & $(1324)^{k-2}132$ & &  $(1324)^{k-2}132$\\
		     & $(3142)^{k-2}3$   & &  $(3142)^{k-2}3$ \vspace{5pt}\\

		     & $(2413)^{k-3}241$ & &  $(2413)^{k-3}241$\\
		     & $(4231)^{k-3}4$   & &  $(4231)^{k-3}4$  \\
		     & $(1324)^{k-4}132$ & &  $(1324)^{k-4}132$\\
		     & $(3142)^{k-4}3$   & &  $(3142)^{k-4}3$\\
			 & \vdots			 & &  \vdots\\
			 & $(2413)^{2} 241$	 & &  $(2413)^{1} 241$\\
			 & $(4231)^2 4$		 & &  $(4231)^1 4$\\
			 & $(1324)^1 132$	 & &  $(1324)^0 132$\\
			 & $(3142)^1 3$		 & &  $(3142)^0 3$\\
			 & $(2413)^0 241$	 & & \\
			 & $(4231)^0 4$  	 & 
	\end{tabular}

\end{center}
The transcript above is written in rows which correspond to beads in the conformation stabilized along the same row on the grid. To simplify the argument we will use \textit{row} both for the transcript above and for the conformation it stabilizes in. In line with this, row $\ell\in \{1,\dots, 2k\}$ is as follows:  
\begin{center}
	\begin{tabular}{llll}
							& row $\ell$ & 			& $\ell\mod 4$ \\
			& $(2413)^{k-\lfloor (\ell+1)/2\rfloor}241$ & \hspace{1cm} &  $\ell\equiv 1\mod 4$\\
							& $(4231)^{k-\lfloor (\ell+1)/2\rfloor}4$   & &  $\ell\equiv 2 \mod 4$ \\
							& $(1324)^{k-\lfloor (\ell+1)/2\rfloor}132$ & &  $\ell\equiv 3 \mod 4$\\
							& $(3142)^{k-\lfloor (\ell+1)/2\rfloor}3$   & &  $\ell\equiv 4 \mod 4$ 
	\end{tabular}
\end{center}
so $w=\mathrm{row}_1\cdots \mathrm{row}_{2k}$. Row $1$ is of length $4k-1$ and row $\ell+1$ is two beads shorter than row $\ell$, so the length of the whole transcript is $|w|=4k^2=\frac{(4k+1-1)^2}{4}=\frac{(|\sigma|-1)^2}{4}$. As an example, see Fig.~\ref{CI:big}, where $k=5$, so the length of the seed is $4k+1=21$ and the transcript is $4k^2 = 100$ beads long.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{./Fig/chichenBig}
	\caption{Quadratic length transcript folding deterministically into pyramid shape. Seed: thick black path. Transcript: thin blue path. Bonds: dashed red lines. Beads: '$0$' black, '$1$' red, '$2$' red with black contour, '$3$' green, '$4$' green with black contour.}
	\label{CI:big}
\end{figure}

Stabilizing the first bead of row $j$ goes as follows (see Fig.~\ref{CI:turn}):
%\begin{itemize}
%	\item $\bf j\equiv 1\mod 4$: the first two beads are $24$, and the preceding bead is stabilized at $(4k-\frac{j-1}{2}, j-1)$. The $2$ can only bind to the $2$ which occurred two beads before at $(4k-\frac{j+1}{2}, j-1)$, and $4$ cannot bind anywhere, so the $2$ is stabilized at $(4k-\frac{j-1}{2},j)$.
%\end{itemize}

\begin{center}
	\begin{tabular}{|c|c|c|c|c|@{}m{0pt}@{}}
		%\renewcommand{\arraystretch}{2}
	$j\mod 4$					& first bead 	& predecessor at		& first bead binds to		& first stabilizes at\\
						\hline
	$\equiv 1$	& 2		& $(4k-\frac{j-1}{2}, j-1)$		& $(4k-\frac{j+1}{2}, j-1)$	& $(4k-\frac{j+1}{2}, j)$& $\mathrm{ }$\newline $\mathrm{ }$\\
	\hline
	$\equiv 2$	& 4				& $(\frac{3j}{2}-1,j-1)$		& $(\frac{3j}{2},j-1)$		& $(\frac{3j}{2},j)$ & $\mathrm{ }$\newline $\mathrm{ }$\\
	\hline
	$\equiv 3$	& 1				& $(4k-\frac{j-1}{2}, j-1)$		& $(4k-\frac{j+1}{2}, j-1)$	& $(4k-\frac{j+1}{2}, j)$ & $\mathrm{ }$\newline $\mathrm{ }$ \\
	\hline
	$\equiv 4$	& 3				& $(\frac{3j}{2}-1,j-1)$		& $(\frac{3j}{2},j-1)$		& $(\frac{3j}{2},j)$ & $\mathrm{ }$\newline $\mathrm{ }$ \\
	\hline		
	\end{tabular}
\end{center}

%The first bead of the transcript is type $2$ and it has to be fixed next to $(4k,0)$. The second bead is type $4$, which cannot form a bond with anything in the conformation so far. The only possibility of forming a bond is if the $2$-bead is placed at $(4k,1)$ and it bonds with the $2$-bead of the seed at $(4k-1,0)$. The next to be stabilized is a $4$-bead, and since it cannot form any bond, the conformation with the highest number of bonds is when the one after, a $1$-bead, is at $(4k-2,1)$ bonding with $(4k-3,0)$, which stabilizes the preceding $4$-bead at $(4k-1,1)$ (see Fig.~\ref{CI:1-4}, (a)-(b)).

\begin{figure}
	\begin{minipage}{.35\textwidth}
		\centering
		\includegraphics[align=c,height=1in]{./Fig/CI_1}\\
		\bigskip

		(a) First transcript bead
	\end{minipage}%
	\begin{minipage}{.2\textwidth}
		\centering
		\includegraphics[align=c,height=0.7in]{./Fig/CI_2}\\
		\bigskip

		(b) Second bead
	\end{minipage}
	\begin{minipage}{.4\textwidth}
		\centering
		\includegraphics[align=c,height=0.5in]{./Fig/CI_3}
		\bigskip
		\vspace{0.1in}
		\includegraphics[align=c,height=0.5in]{./Fig/CI_4}
		\bigskip

		(c) Bead $i+1$, when bead $i$ is $B_{3,4}$ and $B_{1,2}$, respectively
	\end{minipage}
	\caption{Fixing transcript beads in odd rows}
	\label{CI:1-4}
\end{figure}

As for the other beads, in rows $j\equiv 1,3\mod 4$, beads of type $1$ and $3$ bind to a bead in row $j-1$. In rows $j\equiv 2,4\mod 4$ beads of type $2$ and $4$ can bind to a bead in row $j-1$.  This is true for row $1$, because beads of type $1$ and $3$ from row $1$ can only bind to every second bead of the seed, whereas beads of type $2$ and $4$ from row $1$ cannot bind to anything (Fig.~\ref{CI:1-4}, (c)). Once this dynamic holds for a row, it holds inductively for the next, as a bead that binds to another loses its only free hand at arity $1$.

Within one row of the transcript, no bead can bind to a preceding bead, because if there is a previous bead of the same type in that row, it is stabilized at a distance of four and the system has delay $2$.

By the arguments above, the beads in row $i$ of the transcript are stabilized along row $i$ on the grid, forming the pyramid-like conformation from Fig.~\ref{CI:big}.

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{./Fig/CI_turn}
	\caption{Beads $4k, 8k-2, \dots$ stabilize at turning points because the bead two positions before is the same type and has a free hand.}
	\label{CI:turn}
\end{figure}

\subsection{$\alpha = 1$, unary}

Let the point where the first transcript bead was fixed be $p$ and let $n=|\mathrm{seed}|+1$. We will argue about the situation when the first bead is stabilized outside $\hexagon_p^n$ (a hexagon of radius $n$). Let this be the $i$th bead of the transcript. Without loss of generality, we can translate the origin $(0,0)$ to the coordinates of bead $i-1$ (which is still in $\hexagon_p^n$), and we can assume that the bead outside the hexagon is fixed at $(1,1)$ (see Fig.~\ref{fig:hexagonOut}).
\begin{figure}
	\centering
	\includegraphics[width=0.3\linewidth]{./Fig/hexagonOut}
	\caption{$\hexagon_p^n$ and the position $(1,1)$ of the first bead fixed outside of it.}
	\label{fig:hexagonOut}
\end{figure}

In the elongation that places bead $i$ at $(1,1)$ there are two possibilities:
\begin{itemize}
	\item $i$ forms a bond with a bead at $(1,0)$.
	\item  $i$ does not bond to anything and $i+1$ is at $(2,1)$ bonding with a bead at $(2,0)$. If there is no bead at $(1,0)$, then placing $i$ at $(1,0)$ instead of $(1,1)$ results in the same number of bonds, leading to nondeterminism. Therefore, there has to be a bead at $(1,0)$ and it is inactive, otherwise it would bond to $i$. This is analogous to case 1. below.%as in Fig.~\ref{fig:hexagonOut1}.
	\begin{figure}
		\centering
		\includegraphics[width=0.2\linewidth]{./Fig/hexagonOut1}
		%\caption{}
		\label{fig:hexagonOut1}
	\end{figure}

\end{itemize}

%The only position with which $(1,1)$ can form a bond is $(1,0)$. This means that there is a bead at $(1,0)$, which bonds to bead $i$, otherwise there are other conformations in which beads $i$ and $i+1$ add one bond to the conformation, making the behavior nondeterministic.

The next bead, $i+1$, can be fixed at $(2,1)$ or at $(0,1)$ as all other possibilities result in nondeterministic behavior immediately, so we have two cases.

\begin{enumerate}
	\item bead $i+1$ is fixed at $(2,1)$ and can bond with a bead at $(2,0)$. Now consider bead $i+2$. For $i+1$ to be fixed at $(2,1)$, $i+2$ needs to form a bond somewhere, otherwise $i+2$ could go to $(2,1)$ forming the bond with the bead at $(2,0)$ and there would be two conformations with the maximal $1$ bond. The only possibility is that there is a bead at $(3,0)$ and $i+2$ can bond with it when placed at $(3,1)$. We can apply the same argument inductively: there is some $m\geq 0$ such that grid points $(\ell,0)$ are occupied by active beads, for all $\ell\in \{2,\dots,2+m\}$, and there is no bead at $(3+m,0)$. Such an $m$ exists, and it is not greater than $n$. Then, bead $i+\ell$ is fixed at $(\ell+1,1)$ and bonds with $(\ell+1,0)$. However, bead $i+2+m$ cannot be fixed anywhere, because $i+2+m$ and $i+3+m$ can only add one bond to the conformation, and that is possible either with $i+2+m \rightarrow (2+m,1)$, $i+3+m \rightarrow (3+m,1)$ or with $i+2+m \rightarrow (2+m,2)$, $i+3+m \rightarrow (2+m,1)$. 
	\begin{figure}
		\centering
		\includegraphics[width=0.6\linewidth]{./Fig/hexagonOut2}
		\caption{}
		\label{fig:hexagonout2}
	\end{figure}

	\item bead $i+1$ is fixed at $(0,1)$. This is only possible if
	\begin{enumerate}
		\item there is an inactive bead at $(-1,0)$ and an active one at $(-2,0)$. This case is symmetrical to (1).
		\item there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ has two bonds, just as the original conformation.
		\item there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

	\end{enumerate}
\end{enumerate}

\begin{figure}
	\centering
	%\includegraphics[width=0.3\linewidth]{./hexagonOut1}
	%\hspace{10mm} %
	\includegraphics[width=0.6\linewidth]{./Fig/hexagonOut3}

	\caption{}
	\label{fig:hexagonOut2}
\end{figure}

\subsection{In the case of $\delta = 3$}

As before, the length of seed is $n$ and the seed will be inside $\hexagon_p^n$, the equilateral hexagon with a radius of $n$ and center $p$, the point where the first bead of the seed is fixed. We will begin our discussion by considering the moment when the first bead is fixed outside $\hexagon_p^n$. Let this bead be the $i$th one.

\textbf{A.} First, let us assume that a fixed bead on a side of $\hexagon_p^n$ has a free hand, and bead $i$ binds to it. If the most stable conformation formed by beads $i, i+1, i+2$ has only two hydrogen bonds, it will be nondeterministic because there are at least two possibilities, see Fig.~\ref{fig:beadi}. Therefore, it needs to make three hydrogen bonds to deterministically stabilize $i$.\\

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
        \draw (60 : 1.0) -- (120 : 1.0);
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \foreach \theta in {60}{
        \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
        \draw (60 : 1.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \foreach \theta in {60}{
        \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted](0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
    \label{fig:beadi}
  \end{center}
\end{figure}

There are the two cases in which beads $i, i+1, i+2$ can have three bonds, see Fig.~\ref{fig:3bonds}. However, if it makes three bonds once, such as in Fig.~\ref{fig:3bonds}, it will need to make three bonds forever to be deterministic. Similarly to case 1. in the previous section, this becomes nondeterministic finally, as in Fig.~\ref{fig:hexcorner}, when it reaches a corner of $\hexagon_p^n$. Thus, the fixed bead on the side of an equilateral hexagon with a radius of $n$ has to make a hydrogen bond with a bead in an equilateral hexagon.

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (3.5, 0.0);
      \fill (3.5, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (3.5, 0.866);
      \fill (3.5, 0.866) node [right] {$n+1$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \fill (0,0) circle [radius = 0.05];
        \foreach \theta in {0, 60, 120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
          \draw (0 : 0.0) -- (0 : 2.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 3.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-1.5, 0.0) -- (2.5, 0.0);
      \fill (2.5, 0.0) node [right] {$n$};
      \draw [dashed] (-1.5, 0.866) -- (2.5, 0.866);
      \fill (2.5, 0.866) node [right] {$n+1$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (0 : 0.0)]
        \foreach \theta in {0, 60, 120, 180}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \fill (0,0) circle [radius = 0.05];
        \foreach \theta in {120}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
          \draw (0 : 0.0) -- (0 : 2.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (180 : 1.0)]
          \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
    \label{fig:3bonds}
  \end{center}

\end{figure}

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (2.0, 0.0);
      \fill (3.0, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (2.5, 0.866);
      \fill (3.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-0.5, 1.732) -- (3.0, 1.732);
      \fill (3.0, 1.732) node [right] {$n+2$};
      \fill (60 : 1.0) circle [radius = 0.05];
      \draw (-0.1,-0.1) -- (0.1,0.1);
      \draw (0.1,-0.1) -- (-0.1,0.1);
      \begin{scope}[shift = (120 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (300 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \draw (0.0, 0.0) -- (2.0, 0.0);
        \fill (2.0, 0.0) circle [radius = 0.05];
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dashed] (0 : 0.0) -- (300 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
      \fill (0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \fill (0.0, 0.0) circle [radius = 0.05];
          \draw (0 : 0.0) -- (240 : 1.0);
          \draw (0 : 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \draw [dashed] (-0.5, 0.0) -- (2.0, 0.0);
      \fill (3.0, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (2.5, 0.866);
      \fill (3.0, 0.866) node [right] {$n+1$};
      \draw (-0.1,-0.1) -- (0.1,0.1);
      \draw (0.1,-0.1) -- (-0.1,0.1);
      \fill (60 : 1.0) circle [radius = 0.05];
      \begin{scope}[shift = (120 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift = (120 : 1.0)]
        \draw [dotted] (0 : 0.0) -- (300 : 1.0);
        \draw (0.0, 0.0) -- (3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dashed] (0 : 0.0) -- (300 : 1.0);
        \fill (60 : 1.0) circle [radius = 0.05];
      \end{scope}
      \begin{scope}[shift = (0 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {0, 60}{
          \fill (0,0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
    \label{fig:hexcorner}
  \end{center}
\end{figure}

\textbf{B.} Next, assume that bead $i$ could not bind to anything and let us discuss the moment after bead $i$ is fixed outside $\hexagon_p^n$. Now bead $i$ has a free hand. If beads $i+1, i+2, i+3$ can form only two bonds, it will be nondeterministic because there are at least two possible such structures, as in Fig.~\ref{fig:afterzigzag}. Hence, they need to form three bonds to deterministically stabilize, such as in the right or left side of Fig.~\ref{fig:afterline}.\\

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
        \begin{scope}[shift=(0 : 0.0)]
          \draw (-0.1,-0.1) -- (0.1,0.1);
          \draw (0.1,-0.1) -- (-0.1,0.1);
          \draw (0 : 0.0) -- (60 : 1.0);
        \end{scope}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \draw [dashed] (-1.0, 2.598) -- (2.0, 2.598);
      \fill (2.0, 2.598) node [right] {$n+3$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 2.0)]
        \fill(0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {120, 180, 240}{
          \fill (0.0, 0.0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (180 : 1.0);
        \draw (0 : 0.0) -- (240 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \draw (0.0, 0.0) -- (60 : 1.0);
          \draw [dotted] (0.0, 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
        \begin{scope}[shift=(0 : 0.0)]
          \draw (-0.1,-0.1) -- (0.1,0.1);
          \draw (0.1,-0.1) -- (-0.1,0.1);
          \draw (0 : 0.0) -- (60 : 1.0);
        \end{scope}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \draw [dashed] (-1.0, 1.732) -- (2.0, 1.732);
      \fill (2.0, 1.732) node [right] {$n+2$};
      \draw [dashed] (-1.0, 2.598) -- (2.0, 2.598);
      \fill (2.0, 2.598) node [right] {$n+3$};
      \fill (0,0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 2.0)]
        \fill(0.0, 0.0) circle [radius = 0.05];
        \foreach \theta in {120, 180, 240}{
          \fill (0.0, 0.0) [transform canvas = {shift = (\theta : 1.0)}] circle [radius = 0.05];
        }
        \draw (0 : 0.0) -- (120 : 1.0);
        \draw (0 : 0.0) -- (180 : 1.0);
        \draw [dotted] (0 : 0.0) -- (240 : 1.0);
      \end{scope}
      \begin{scope}[shift = (60 : 1.0)]
        \begin{scope}[shift = (120 : 1.0)]
          \draw [dotted] (0.0, 0.0) -- (60 : 1.0);
          \draw (0.0, 0.0) -- (300 : 1.0);
        \end{scope}
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
    \label{fig:afterzigzag}
  \end{center}
\end{figure}

 If they form three bonds once, such as Fig.~\ref{fig:afterline}, the system would need three bonds forever to be deterministic. This becomes nondeterministic finally, such as in Fig.~\ref{fig:hexcorner}, when it reaches a corner of $\hexagon_p^n$. 
 %Therefore, the first fixed bead on a side of an equilateral hexagon with a radius of $n+1$ has to make a hydrogen bond with a molecule on a side of an equilateral hexagon with a radius of $n$.

\begin{figure}
  \begin{center}
  \begin{tabular}{cc}
    \begin{tikzpicture}
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \draw [dashed] (-0.5, 0.0) -- (4.5, 0.0);
      \fill (4.5, 0.0) node [right] {$n$};
      \draw [dashed] (-0.5, 0.866) -- (4.5, 0.866);
      \fill (4.5, 0.866) node [right] {$n+1$};
      \fill (2.0, 0.0) circle [radius = 0.05];
      \fill (3.0, 0.0) circle [radius = 0.05];
      \fill (4.0, 0.0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \fill (1.0, 0.0) circle [radius = 0.05];
        \fill (2.0, 0.0) circle [radius = 0.05];
        \fill (3.0, 0.0) circle [radius = 0.05];
        \draw (0.0, 0.0) -- (3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : 2.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 3.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : 4.0)]
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}

    \begin{tikzpicture}
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \draw [dashed] (-3.5, 0.0) -- (1.0, 0.0);
      \fill (1.0, 0.0) node [right] {$n$};
      \draw [dashed] (-3.5, 0.866) -- (1.0, 0.866);
      \fill (1.0, 0.866) node [right] {$n+1$};
      \fill (-1.0, 0.0) circle [radius = 0.05];
      \fill (-2.0, 0.0) circle [radius = 0.05];
      \fill (-3.0, 0.0) circle [radius = 0.05];
      \begin{scope}[shift = (60 : 1.0)]
        \fill (0.0, 0.0) circle [radius = 0.05];
        \fill (-1.0, 0.0) circle [radius = 0.05];
        \fill (-2.0, 0.0) circle [radius = 0.05];
        \fill (-3.0, 0.0) circle [radius = 0.05];
        \draw (0.0, 0.0) -- (-3.0, 0.0);
      \end{scope}
      \begin{scope}[shift = (0 : -1.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : -2.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift = (0 : -3.0)]
        \draw [dotted] (0 : 0.0) -- (60 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \end{tabular}
    \caption{}
    \label{fig:afterline}
  \end{center}
\end{figure}

Accordingly, it is Fig.~\ref{fig:firstout} when a first bead is fixed on a side of an equilateral hexagon with a radius of $n+1$.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw [dashed] (-1.0, 0.0) -- (2.0, 0.0);
      \fill (2.0, 0.0) node [right] {$n$};
      \draw [dashed] (-1.0, 0.866) -- (2.0, 0.866);
      \fill (2.0, 0.866) node [right] {$n+1$};
      \begin{scope}[shift=(0 : 0.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw (0 : 0.0) -- (60 : 1.0);
      \end{scope}
      \begin{scope}[shift=(60 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
      \end{scope}
      \begin{scope}[shift=(0 : 1.0)]
        \draw (-0.1,-0.1) -- (0.1,0.1);
        \draw (0.1,-0.1) -- (-0.1,0.1);
        \draw [dotted] (0 : 0.0) -- (120 : 1.0);
      \end{scope}
    \end{tikzpicture}
    \caption{}
    \label{fig:firstout}
  \end{center}
\end{figure}

%It has to be a right or left figure in Fig.6. when the first bead fixed on a side of an equilateral hexagon with a radius of $n+1$ make a hydrogen bond with a bead on a side of an equilateral hexagon with a radius of $n$. Fig.6. becomes nondeterministic finally, such as in Fig.7., when it reaches a corner of an equilateral hexagon. It is clear that all beads on a side of an equilateral hexagon with a radius of $n+1$ has to make a hydrogen bond with a bead on a side of an equilateral hexagon with a radius of $n$.  It is impossible to fix molecules on a side of an equilateral hexagon with a radius of $n+2$ and determine the only structure when there are only beads which are able to make a hydrogen bond in an equilateral hexagon with a radius of $n$.