### 2018-12-05 tt_theorem.tex
12/5 morning

%remark

\begin{theorem}[Tunnel Troll Theorem]
\label{TTT}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$. The following statements hold.
\begin{enumerate}
\item At arity $\alpha \geq 3$, if $S[i] = t$ (i.e. the tunnel that stabilizes $w[i]$ is not singular) and $S[i+1] \neq \blacksquare$, then $\#bc(C_{i-1}) > \#bc(C_i)$.
\item At arity $\alpha = 2$, if $n$ is the number of $bt$ in $S[i..j]$ for some $i,j$ such that $S[i] = b$ and $i < j$, then $\#bc(C_{i-2}) - \#bc(C_{j}) \geq n$.
\end{enumerate}
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[tb]
  \begin{center}
    \begin{tabular}{ccc}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
        \draw[->, white] (0:0)--(-120:1);
	 \draw[->,white] (120:0.8)--(120:0.1);%for adjustment

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)},blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {120,-120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (180:0.9)--(180:0.1);
            \draw[->, blue] (0:0.1)--(0:0.9);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};

         % \node at (0,-2) {$t_{0}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
	\draw[->, white] (0:0)--(-120:1);
            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)},blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {180,-120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (120:0.9)--(120:0.1);
            \draw[->, blue] (0:0.1)--(0:0.9);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          %\node at (0,-2) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
        \draw[->, white] (0:0)--(-120:1);
            \fill[blue](0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

            \foreach \theta in {0}{
              \fill[transform canvas={shift=(\theta:1)}, blue](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {180,-120,120}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

            \draw[->, blue] (60:1.1)--(60:0.1);
            \draw[->, blue] (0:0.1)--(0:1.1);

            %\node[transform canvas={shift=(0:1)},below] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

         % \node at (0,-2) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{Direction into a entrance}
    \label{TTT_tunnel_direction}
  \end{center}
\end{figure}

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be a deterministic unary oritatami system at delay $\delta = 1$, arity $\alpha =2$. 
Assume $\Xi$ stabilizes the transcript until $w[i-1]$. If $S[i+1] = b$ and $S[i+2] \in \{ t_s, t_o \} $, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

\begin{proof}%[lemma \ref{TTT_entrance_Tab}]
See Fig.\ref{TTT_tunnel_direction}.
$S[i+2] \in \{ t_s, t_o \} $ means that $w[i+2]$ is stabilized by a tunnel section of type $S$ or $O$.
Thus,  its predecessor $w[i+1]$ must be inside the tunnel section, that is, $n_1$ and $n_2$ must be occupied.
Free bonds of $w[i]$, if any, cannot be used in future by another bead $w[j]$ because otherwise the part of transcript $w[i..j]$ and the bond between $w[i]$ and $w[j]$ would form a closed curve and the curve would cross the path of $C_{i-1}$ between $n_1$ and $n_2$, contradiction.
%Binding capabilities that $w[i]$ supplies are inactive according to Jordan curve theorem on $n_1$, $n_2$, and $w[i]$.
Therefore, if $w[i]$ forms a bond at its stabilization $\#bc(C_{i-1}) > \#bc(C_{i})$ holds.
We now prove that $w[i]$ must form a bond.

Suppose $w[i]$ were stabilized without any bond, that is, by a tunnel.
For that the two points that are a neighbor of both $w[i-1]$ and $w[i]$ must be occupied already.
In addition, at least one of the neighbors of $w[i]$ must be free because $S[i+1] = b$.
Thus, only the case to be considered is Fig.~\ref{TTT_tunnel_direction} (middle) with $n_5$ being occupied (that is, $n_4$ is free).
In this case, before $w[i]$ is stabilized, at lest three neighbors of $n_2$ were free and hence, a bead at $n_2$ was provided with one free bond and could form a bond with $w[i]$. \qed

\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]
  \centering
    \begin{tabular}{cc}
    \centering
      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-60:1);
	\draw[white] (0:0) -- (0:1);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$w[j]$};

            \foreach \theta in {120,-120,180}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {0}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }
            \draw (-60:1) circle [radius=0.05];
            \draw (60:1) circle [radius=0.05];

            \draw[->, blue] (180:0.9)--(180:0.1);

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},above] {$w[j-1]$};

          \node at (0,-2) {Point $n_1$ and $n_2$ are free};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.45\linewidth}
      \centering
        \begin{tikzpicture}
	\draw[white] (0:0) -- (0:1);
	\draw[white] (0:0) -- (120:1);
	\draw[white] (0:0) -- (-90:2);

            \fill[blue](0,0) circle [radius=0.05];
            \node[above] at (0,0) {$w[j]$};

            \foreach \theta in {120,-120,180,60}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }

            \foreach \theta in {0}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }
            \draw (-60:1) circle [radius=0.05];

            \draw[->, blue] (180:0.9)--(180:0.1);

            \node[transform canvas={shift=(0:1)},right] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},above] {$w[j-1]$};

          \node at (0,-2) {Point $n_1$ is occupied};
        \end{tikzpicture}
      \end{minipage}
    \end{tabular}
    \caption{Exit of Tunnel}
    \label{TTT_tunnel_exit}
\end{figure}

\begin{lemma}
\label{TTT_exit}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$ and arity $\alpha =2$.
%Let $w[i]$ be a bead which is stabilized at the exit of a tunnel.
If $S[i+1..j+1] = bt^{(j-i-1)}b$ for some $i,j$ with $i \leq j-2$ and $S[j] \in \{ t_s, t_o\}$, then $\#bc(C_{j-2}) \geq \#bc(C_j)$, and hence, $\#bc(C_{i}) \geq \#bc(C_j)$.
If $i \leq j-3$, then the second inequality is strengthened as $\#bc(C_{i}) > \#bc(C_{j})$.
%if we assume $\#bc(C_{i-2}) - \#bc(C_{i-1}) = a$, then $\#bc(C_{i}) - \#bc(C_{i-1}) \leq a$.
\end{lemma}

\begin{proof}%[Lemma~ \ref{TTT_exit}]
Since the binding capability never increases inside a tunnel, we just need to consider the exit of a tunnel.
See Fig.\ref{TTT_tunnel_exit}. 
At least one of points $n_1$ or $n_2$ must be free because otherwise $w[j]$ would be inside of a tunnel, that is, $S[j+1]$ would not be $b$.

%If $i - h > 1$ and $j$ is such that $h \leq j<i$, then $\#bc(C_{j-1}) \geq \#bc(C_{j})$ because all neighbors of $w[j]$ are occupied by beads forming the tunnel so that any $w[i+1..]$ cannot reach neighbors of $w[j]$. 
%Thus, inside this tunnel, binding capabilities never increased.

Let $m$ be the number of bonds $w[j-1]$ forms, that is, $\#bc(C_{j-2}) - \#bc(C_{j-1}) = m$.
We claim $\#bc(C_{j}) - \#bc(C_{j-1}) \leq m$.
Indeed, if both $n_1$ and $n_2$ are free (see Fig.~\ref{TTT_tunnel_exit}), the predecessor $w[j-1]$ must be bound to both beads at $n_3$ and at $n_4$ because both of them still have a free hand.
Hence, $m \geq 2$.
Since $\#bc(C_{j}) - \#bc(C_{j-1})$ is less than the arity, this difference is at most $m$.

If $n_1$ is occupied, then $n_2$ is free.
The predecessor $w[j-1]$ must be bound $n_4$.
Hence,  $m \geq 1$.
The bead $w[j]$ can increase the binding capability at most by 1 because one of its free neighbors would, $n_0$ or $n_2$, is to be occupied by the successor $w[j+1]$.
Therefore, $\#bc(C_{j}) - \#bc(C_{j-1}) \leq m$.

Thus,$\#bc(C_{j-2}) \geq \#bc(C_j)$ , and hence, $\#bc(C_{i}) \geq \#bc(C_j)$.
If $i \leq j-3$, then the second inequality is strengthened as $\#bc(C_{i}) > \#bc(C_{j})$ because $S[i+1] = b$, that is, $\#bc(C_{i}) > \#bc(C_{i+1})$ and $\#bc(C_{i+1}) \geq \#bc(C_j)$. \qed

\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \begin{center}
        \begin{tikzpicture}
        \draw[->,white] (-70:0.1) -- (-70:3);
          \foreach \theta in {60,-60,120,0}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

          \fill[blue](180:1) circle [radius=0.05];
          \fill[blue](0:0) circle [radius=0.05];

          \draw[transform canvas={shift=(180:1)}, blue] (105:1) edge[bend right] (105:0);
          \draw[->, blue] (180:0.9) -- (180:0.1);
          \draw[->, blue] (-120:0.1) -- (-120:0.9);

          \node[below left] at (180:1) {$w[i]$};
          \node[above] at (0:0) {$w[i+1]$};
          \node[below] at (-120:1) {$w[i+2]$};

          \node[right] at (120:1) {$n_3$};
          \node[right] at (60:1) {$n_2$};
          \node[right] at (0:1) {$n_1$};

          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];
            \foreach \theta in {-120,-60,180}{
              \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
            }

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[h-1..h+1] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

\begin{lemma}
\label{TTT_tunnelC_lemma}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$, arity $\alpha = 2$.
If $S[i+2] = t_a$, the following statements hold.
\begin{enumerate}
\item If $w[i]$ forms at least one bond, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consumes any bond and $S[i] \in \{t_s, t_o \}$, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consumes any bond and $S[i] = t_a$, $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
\end{enumerate}
%If there are indices $i,j$ such that $S[i..j+1]$ is either $bbt^{(j-i-1)}b$ and $S[i+2] = t_a$ or $bt^\ell bt^{j-i-1-\ell}b$ for some $\ell$ and $S[i+l+2] = t_a$, then $\#bc(C_{i-1}) > \#bc(C_j)$.
\end{lemma}

\begin{proof}
%	The bead $w[i+1]$ consumes one hand and provide nothing, that is, $\#bc(C_{i}) \geq \#bc(C_{i+1})$.
%	The bead $w[i]$ consumes bond because $S[i] = b$.
%	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
%	The bead $w[i+2]$ is stabilized at a point of neighbors of $w[i]$.
%	If $w[i]$ consumes just one hand, $w[i]$ leaves a hand, that is, $\#bc(C_{i-1}) \geq \#bc(C_{i})$.
%	Then, $w[i+2]$ consumes $w[i]$'s hand.
%	Thus, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
%	On the other hand, if $w[i]$ forms two bonds, then even if $w[i+2]$ provides two free hands, $\#bc(C_{i-1}) - \#bc(C_{i}) = \#bc(C_{i+1}) - \#bc(C_{i+2})$.
%	Thus, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
%	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	We consider each statements. First we prove statement 1.
	The bead $w[i+1]$ consumes one hand and provide nothing.
	If $w[i]$ forms two bonds, then even if $w[i+2]$ provides two free hands, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	On the other hand, if it leaves a free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Next we prove statement 2.
	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $S$ or $O$.
	As prove in Lemma~\ref{TTT_exit}, $\#bc(C_{i-2}) \geq \#bc(C_{i})$.%the increase of binding capability by $w[i-1]$ and $w[i]$ is at most 0.
	The bead $w[i]$ leaves two free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Finally, we prove statement 3.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $A$ and $w[i]$ does not consumes any bond.
	Then, $w[i-2]$ must consume two bonds; otherwise $w[i-2]$'s hands is used by $w[i]$.
	The bead $w[i-1]$ consumes one hand and provide nothing.
	Thus,  $\#bc(C_{i-3}) - 1 \geq \#bc(C_{i})$.
	Moreover, the bead $w[i+1]$ consumes one hand and provide nothing and $w[i]$'s hands is used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Therefore, $\#bc(C_{i-3}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.
\end{proof}

%\begin{lemma}
%\label{TTT_tunnelC_lemma}
%Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$, arity $\alpha = 2$. We assume $S[h..i+1] = bt^{(i-h)}b \quad (1<h<i)$. If at least one of $w[h+1..i]$ is stabilized by tunnel $C$, then $\#bc(C_{h-3}) > \#bc(C_{h+1})$ and $\#bc(C_{h-3}) > \#bc(C_i)$.
%\end{lemma}

%\begin{proof}[Lemma~ \ref{TTT_tunnelC_lemma}]
%Let $\Xi$ be a unary oritatami system at $\delta = 1, \alpha = 2$.
%Assume $S[h..i+1] = bt^{i-h}b (h<i)$. If at least one of $w[h+1..i]$ are stabilized by tunnel $C$, then only $w[h+1]$ can use tunnel $C$ because if $w[g]$ which is one of $w[h+2..i]$, with $h+2 \leq g \leq i$ is stabilized by tunnel $C$, $C_g$ is a terminal.
%
%
%Let us consider stabilization $S[h-1..h+1] = tbt$ or $S[h-1..h+1] = bbt$ as follows. In result, $\#bc(C_{h-3}) > \#bc(C_{h+1})$. In addition according to Lemma~\ref{TTT_exit} $\#bc(C_{h+1}) \geq \#bc(C_{i})$. Thus, $\#bc(C_{h-3}) > \#bc(C_{h+1})$ and $\#bc(C_{h-3}) > \#bc(C_{i})$.
%\\
%
%\subsubsection{Case of $S[h-1..h+1] = bbt$}
%See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
%If $w[h-1]$ forms one bond, then it leaves one free bond, but it is consumed by $w[h+1]$.
%In addition, $w[h]$ has to bound.
%Thus, in this cases, binding capability decreases by 1.
%If $w[h-1]$ forms two bonds, then it does not leave any bond.
%In addition, $w[h]$ has to be bound.
%Thus, in this case, even if $w[h+1]$ supplies two free bonds, binding capability decreases by 1.

%\subsubsection{Case of $S[h-1..h+1] = tbt_a$}
%Fig.\ref{TTT_tunnelC_enter_usingTunnel} exhibits all the two kinds of stabilization depending on structures of tunnel C.
%
%%\item{Left of Fig.\ref{TTT_tunnelC_enter_usingTunnel}}\\
%  In this figure, Bead $n_4$ has at least one binding so that $w[h-1]$ has to bound $n_4$. Moreover, $w[h]$ has to bind to one of $n_1, n_2, n_3$ in order to stabilize deterministically. On the other hand, $w[h+1]$ can supply two bindings but has only two free neighbors. One of them is occupied by a successor. Therefore $w[h+1]$ can only bind one of $n_5, n_6$, that is, $w[h+1]$ supplies at most one binding. Thus, this case $\#bc(C_{h-1}) > \#bc(C_{h+1})$.
%
%  These cases are divided on number of capabilities that $w[h-1]$ consumes.
%  \begin{itemize}
%  \item[-]{$w[h-1]$ does not consume any bindings}\\
%  According to Lemma~\ref{TTT_exit}, $\#bc(C_{h-3}) \geq \#bc(C_{h-1})$ because of $S[h-1] = t$.
%    $w[h]$ has to bound one of $n_1, n_2, n_3$ in order to stabilize deterministically so that  $\#bc(C_{h-1}) > \#bc(C_{h})$.
%    $w[h+1]$ has to be bound to $w[h-1]$ because $w[h-1]$ has bindings, that is, $w[h+1]$ consumes at least one hand and supplies at most one hand so that $\#bc(C_{h}) \geq \#bc(C_{h+1})$. Thus, in this cases $\#bc(C_{h-3}) > \#bc(C_{h+1})$.
%%     This time, let us consider either $n_4$ is occupied or not. If $n_4$ is occupied, then $w[h-1]$ has no active bindings that is this situation consumes some binding capabilities. If $n_4$ is free and $w[h+2]$ is stabilized in $n_4$, then $w[h-1]$ has to bind $w[h+2]$. Therefore, In this case, stabilization of $w[h-1..h+2]$ consumes some bindings. If $n_4$ is free and $w[h+2]$ is stabilized except $n_4$, then this oritatami system has to use two binding capabilities in order to bind $w[h+2]$. Therefore, in this case consumes some bindings. Thus in this cases $\#bc(C_{h-1}) > \#bc(C_{h+2})$.
%
%  \item[-]{$w[h-1]$ consumes one binding}\\
%    In this case, $w_{h}$ has to be bound one of $n_1, n_2, n_3$. In addition, $w[h-1]$ and $w[h+1]$ are not supply any bindings. Thus, in this cases consume some binding capabilities.
%  \item[-]{$w[h-1]$ consumes two bindings}\\
%    In this case, $w[h-1]$ already consumes two binding. $w[h]$ has to be bound. $w[h+1]$ supplies two bindings. Thus, in this cases $\#bc(C_{h-1}) > \#bc(C_{h+1})$.
%   
%  \end{itemize}

%\begin{figure}
%  \centering
%    \begin{tabular}{cc}
%      
%      \begin{minipage}{0.48\hsize}
%      \centering
%        \begin{tikzpicture}
%
%          \fill[transform canvas={shift=(0:0)}](0,0) circle [radius=0.1];
%          
%          \foreach \theta in {60,-60,120,-120,180}{
%            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
%          }
%
%          \fill[blue](0:1) circle [radius=0.05];
%          \draw[->, blue] (0:0.1)--(0:0.9);
%
%          \node[below] at (-60:1) {$n_4$};
%
%          \begin{scope}[shift=(0:2)]
%            \fill[blue](0,0) circle [radius=0.05];
%
%            
%            \foreach \theta in {120,60,-60,0}{
%             \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
%            }
%
%            \draw[->, blue] (180:0.9)--(180:0.1);
%            \draw[->, blue] (-120:0.1)--(-120:0.9);
%
%            \node[below] at (180:1) {$w_{[h-1]}$};
%            \node[above] at (0:0) {$w_{[h]}$};
%            \node[below] at (-120:1) {$w_{[h+1]}$};
%
%            \node[above] at (120:1) {$n_3$};
%            \node[above] at (60:1) {$n_2$};
%            \node[right] at (0:1) {$n_1$};
%
%            \begin{scope}[shift=(-120:1)]
%              \fill[blue](0,0) circle [radius=0.05];
%              \foreach \theta in {-120,-60}{
%                \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
%              }
%
%              \node[right] at (-120:1) {$n_5$};
%              \node[right] at (-60:1) {$n_6$};
%            \end{scope}
%            
%          \end{scope}
%
%          \node at (1.2,-2.5) {Pattern 1};
%        \end{tikzpicture}
%      \end{minipage}
%
%      \begin{minipage}{0.48\hsize}
%      \centering
%        \begin{tikzpicture}
%
%          \fill(0,0) circle [radius=0.1];
%          
%          \foreach \theta in {60,0,120,-120,180}{
%            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
%          }
%
%          \draw[->, blue] (-60:0.1)--(-60:0.9);
%
%
%          \begin{scope}[shift=(-60:1),shift=(0:1)]
%            \fill[blue](0,0) circle [radius=0.05];
%            \fill[blue](180:1) circle [radius=0.05];
%
%            
%            \foreach \theta in {120,60,0,-60}{
%              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
%            }
%
%            \draw[->, blue] (180:0.9)--(180:0.1);
%            \draw[->, blue] (-120:0.1)--(-120:0.9);
%
%            \node[below] at (180:1) {$w_{[h-1]}$};
%            \node[above] at (0:0) {$w_{[h]}$};
%            \node[below] at (-120:1) {$w_{[h+1]}$};
%
%            \node[above right] at (120:1) {$n_3$};
%            \node[right] at (60:1) {$n_2$};
%            \node[right] at (0:1) {$n_1$};
%
%            \begin{scope}[shift=(-120:1)]
%              \fill(0,0) circle [radius=0.05];
%              \foreach \theta in {180,-120,-60}{
%                \draw[transform canvas={shift=(\theta:1)}](-0.05,-0.05) rectangle (0.05,0.05);
%              }
%
%              \node[left] at (180:1) {$n_4$};
%              \node[right] at (-120:1) {$n_5$};
%              \node[right] at (-60:1) {$n_6$};
%            \end{scope}
%            
%          \end{scope}
%
%          \node at (1.25,-4) {Pattern 2};
%        \end{tikzpicture}
%      \end{minipage}
%
%      
%      
%    \end{tabular}
%    \caption{Case of $S[h-1..h+1] = tbt$}
%    \label{TTT_tunnelC_enter_usingTunnel}
%\end{figure}

%\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%%%
%%%%%%
%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Each bead in the transcript is bound either inside a tunnel or outside. If a bead is stabilized inside a tunnel, 
%then it has at most one free neighbor, and hence its successor it to be stabilized there.
%Moreover, if a bead is stabilized outside a tunnel, then its position is either an entrance of a tunnel or not.
%
%Tunnel sections have three possible shapes up to symmetry : straight($S$), obtuse($O$) and acute($A$) turn (Fig. \ref{fig:TTT_tunnel}), and we will consider each of those.

%\begin{lemma}
%\label{TTT_neighbor_lemma}
%For unary transcripts at $\delta = 1$, if a bead has no free hand, then at least $\alpha + 2$ of its neighbors have to be occpied.
%\end{lemma}

\proof{Proof of Tunnel Troll Theorem}

Let us first consider cases of $\delta \geq 3, \alpha = 1$. 
See Fig.~\ref{TTT_tunnel_exit}. Consider the stabilization of $w[i]$. 
This bead $w[i]$, once stabilized, shares two neighbors with its predecessor $w[i-1]$, which are denoted by $n_3, n_4$. 
Both of them have been already occupied because $S[i] = t$.

Since $S[i+1] \neq \blacksquare$, at least one of the other three neighbors, denoted by $n_0, n_1, n_2$, must be free. 
Assume that in the neighborhood of $w[i]$, there are two beads with one free neighbor even after $w[i]$ is stabilized. 
Before the stabilization of $w[i]$, such a bead had two free neighbors, and hence, is provided with at least one free bond. 
Thus, $w[i]$ is to be bonded to these two beads, and it decreases the binding capability by at least 1. 
It now suffices to check that this assumption holds no matter how $n_0, n_1, n_2$ are occupied as long as at least one of them is left free.

Next, we consider the case of $\delta = 2, \alpha = 1$. We assume there are an indices $i$ ,$j$ such that $S[i+1..j+1] = bt^{(j-i-1)}b$ and all $S[i+2..j]$ are  $ t_s$ or  $t_o$.
According to Lemma~\ref{TTT_entrance_Tab}, $\#bc(C_{i-1}) > \#bc(C_{i})$ and moreover that, according to Lemma~\ref{TTT_exit}, $\#bc(C_{i}) \geq \#bc(C_{j})$.
Thus, binding capability of the conformation decrease per a subsequence $bt_s$ or $bt_o$.

On the other hand, we assume there are an indices $i$ ,$j$ such that $S[i+1..j+1] = bt^{(j-i-1)}b$ and there is a element of $S[i+2..j]$ such that it is $t_a$.
All $S[i+3..j]$ do not contain $t_o$; otherwise $S[i+4..j+1]$ contain $\blacksquare$.
Therefore, $S[i+2] = t_a$.
According to Lemma~\ref{TTT_exit}, on statement 1, 2, $\#bc(C_{i-2}) > \#bc(C_{i+2})$ and if $i \leq j - 3$, $\#bc(C_{i+2}) \geq \#bc(C_{j})$.
Thus, binding capability of the conformation decrease per a subsequence $bt_a$ unless $w[i]$ does not consumes any bond.
If $w[i]$ does not consumes any bond and $S[i] = t_a$, that is, $S[i-1..i+2] = bt_abt_a$, then $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
In this case, two subsequences of $bt_a$ collaborate decreasing binding capability but eventually, two subsequences decrease two bond.
That is, one subsequence decrease binding capability.

Hence, if $S[i..j]$ contain n's subsequences of $bt$, $\#bc(C_{i-2}) - \#bc(C_{j}) \geq n$
\qed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/////////////////////////////////////////////////
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{On structures provided by a unary and $\delta = 1$ oritatami system}

%\begin{figure}[tb]
% \centering
%    \begin{tikzpicture}
%	\fill[shift=(180:1)] (0,0) circle [radius=0.1];
%      \fill[shift=(180:0)] (0,0) circle [radius=0.1];
%      
%      \draw[->] (180:0.9) -- (180:0.1);
%
%	\node[above] at (180:1) {$w_{[i-2]}$};
%	\node[below] at (180:0) {$w_{[i-1]}$};
%	\node[above] at (0:1) {$n_0$};
%
%	\foreach \theta in {0,60,-60,120,-120}{
%   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
%  	}
% 	\node[above] at (60:1) {$n_1$};
%	\node[below] at (-60:1) {$n_2$};
%	\node[above] at (120:1) {$n_3$};
%	\node[below] at (-120:1) {$n_4$};
%    \end{tikzpicture}
%    \caption{$\alpha = 4$: when $w[i]$ is stabilized}
%    \label{TTT_a4_lemma_fig}
%\end{figure}

%\begin{lemma}
%\label{TTT_a4_2b_lemma}
%Let $\Xi$ be a deterministic unary oritatami system of $\delta = 1, \alpha = 4$.
%When a bead is stabilized by bond, it forms at least two bonds.
%\end{lemma}
%
%\begin{proof}%[proof of lemma]
%Suppose $w[i]$ were stabilized by only one bond. See Fig.~\ref{TTT_a4_lemma_fig}. If $n_3$ is free, $w[i-2]$ still has some bonds because of arity $\alpha = 4$. Thus, $n_4$ must be occupied; otherwise, $w[i]$ would be stabilized nondeterministically at $n_3$ and $n_4$. Moreover, also $n_2$ has to be occupied for deterministic and also $n_0, n_1$. $n_1$ has some hands because $n_3$ is free. Therefore, $w[i]$ is stabilized at $n_3$ and it has to use at least two hands. It contradict assumption.
%\end{proof}

\begin{figure}[tb]
 \centering
    \begin{tikzpicture}

	\fill (0,0) circle [radius=0.1];
        \fill[blue] (60:1) circle [radius=0.05];
        \fill (0:1) circle [radius=0.1];

      \draw[dashed] (180:2) -- (0:2);
      \draw[dashed, shift=(60:1)] (180:2.5) -- (0:1.5);
      \draw[->, blue] (60:0.1) -- (60:0.9);

	\node[above] at (180:1) {$p_1$};
	\node[left] at (-120:1) {$p_2$};
	\node[left] at (-60:1) {$p_3$};
	\node[left, shift=(-120:1)] at (-60:1) {$p_3$};

	\draw (180:1) circle [radius=0.05];
	\draw (-120:1) circle [radius=0.05];
	\draw (-60:1) circle [radius=0.05];
	\draw[shift=(-120:1)] (-60:1) circle [radius=0.05];

	\node at (0:2.5) {$n$};
	\node[shift=(60:1)] at (0:2) {$n+1$};
    \end{tikzpicture}
    \caption{the first bead out of $\hexagon_{w[-n+1]}^n$}
    \label{TTT_a4_first}
\end{figure}

\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a deterministic unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only finite conformation of length is $3n^2  + 3n + 1$.
\end{theorem}

\begin{proof}
Consider the moment when a bead $b$ is stabilized outside $\hexagon_O^n$ for the first time. 
See Fig.~\ref{TTT_a4_first}.
There is only one bead available to stabilize $b$ there, which is $b_1$. 
In order for $b$ to avoid nondeterminism, hence, none of the beads around should not attract $b$.

The point $p_1$ must be empty because a bead there would have at least two free neighbors and hence is provided with a free hand. 
If there is a bead at $p_2$, $n$ must be at least 2 so that the bead is not singular. 
Since $p_1$ is empty, this bead has at least one free hand, a contradiction. 
Thus, $p_2$ must be also empty. 
In the same way, we can easily show that the point $p_3$ must not be occupied by a non-singular bead. 
Suppose $p_3 = O$. 
The point $p_4$ must not be empty; otherwise the singular bead, at $O$, would have a free hand. 
However, then a bead at $p_4$ would be provided with a free hand, a contradiction. 
\qed
\end{proof}

%\begin{lemma}
%\label{TTT_a3_2b_lemma}
%Let $\Xi$ be a deterministic unary oritatami system of $\delta = 1, \alpha = 3$. 
%If $S[i] = t$ or $t^\times$ and $S[i+1] \neq \blacksquare$, then the bead $w[i]$ consumes also at least one bond. 
%% Let $p$ be a point whose neighbors is occupied at least two point. If $w[i]$ is not stabilized and $w[i-1]$ includes neighbors of $p$, then $w[i]$ is stabilized at $p$ with at least one bond, $w[i]$ is stabilized at another point of $p$ otherwise with at least two bond except any neighbors of $p$ is occupied.
%\end{lemma}
%
%\begin{proof}%[proof of lemma]
%Both $n_3$ and $n_4$ are occupied because $S[i] = t$ or $t^\times$.
%Suppose $w[i]$ were stabilized without any bond.
%Then, all neighbors of $n_3$ must be occupied as long as the bead at $n_3$ is not singular; otherwise, a bead at $n_3$ would be capable of forming a bond with $w[i]$.
%That is, $n_1$ is occupied.
%We can continue this argument in clockwise manner to fill the neighbors of $w[i]$ until a singular bead is encountered, if any.
%If needed, the same argument is applied starting from $n_4$ in counter-clockwise manner.
%Thus, all neighbors of $w[i]$ must be occupied but then $S[i+1] = \blacksquare$, contradiction. \qed
%\end{proof}

\begin{figure} [tb]
  \begin{center}
  \begin{tabular}{c c c}
 \begin{minipage}{0.3\hsize}
 \centering
    \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](0:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (0:0.1) -- (0:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (0:1) {$w_{[i]}$};

	\foreach \theta in {60,-60,120,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	 \draw [shift=(60:1)] (0.95,-0.05) rectangle (1.05,0.05);
	  \draw [shift=(180:1),shift=(120:1)] (-0.05,-0.05) rectangle (0.05,0.05);
 	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(0:1)] at (60:1) {$n_{-1}$};
	\node[above, shift=(180:1)] at (120:1) {$n_5$};
    \end{tikzpicture}
    \end{minipage}

    \begin{minipage}{0.3\hsize}
    \centering
     \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](60:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (60:0.1) -- (60:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (60:1) {$w_{[i]}$};

	\foreach \theta in {0,-60,120,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	\draw [shift=(60:1)] (0.95,-0.05) rectangle (1.05,0.05);
	\draw [shift=(60:1), shift=(60:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(60:1), shift=(120:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	  \draw [shift=(180:1),shift=(-120:1)] (-0.05,-0.05) rectangle (0.05,0.05);

 	\node[right] at (0:1) {$n_0$};
	\node[below] at (-60:1) {$n_2$};
	\node[above] at (120:1) {$n_3$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(60:1)] at (0:1) {$n_{-1}$};
	\node[above, shift=(60:1)] at (60:1) {$n_{-2}$};
	\node[above, shift=(60:1)] at (120:1) {$n_{-3}$};
	\node[below, shift=(180:1)] at (-120:1) {$n_5$};
    \end{tikzpicture}
    \end{minipage}

    \begin{minipage}{0.3\hsize}
        \centering
     \begin{tikzpicture}

      \fill[shift=(180:1)] (0,0) circle [radius=0.1];
      \fill[shift=(180:0)] (0,0) circle [radius=0.1];

      \fill[blue](120:1) circle [radius=0.05];

      \draw[->] (180:0.9) -- (180:0.1);
      \draw[->, blue] (120:0.1) -- (120:0.9);

	\node[above] at (180:1) {$w_{[i-2]}$};
	\node[below] at (180:0) {$w_{[i-1]}$};
	\node[above] at (120:1) {$w_{[i]}$};

	\foreach \theta in {0,60,-60,-120}{
   	   \draw [shift=(\theta:1)] (-0.05,-0.05) rectangle (0.05,0.05);
  	}
	\draw [shift=(120:1), shift=(60:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(120:1), shift=(120:1)] (-0.05,-0.05) rectangle (0.05,0.05);
	\draw [shift=(120:1), shift=(180:1)] (-0.05,-0.05) rectangle (0.05,0.05);

 	\node[right] at (0:1) {$n_0$};
	\node[above] at (60:1) {$n_1$};
	\node[below] at (-60:1) {$n_2$};
	\node[below] at (-120:1) {$n_4$};
	\node[above, shift=(120:1)] at (60:1) {$n_{-1}$};
	\node[above, shift=(120:1)] at (120:1) {$n_{-2}$};
	\node[above, shift=(120:1)] at (180:1) {$n_{-3}$};
    \end{tikzpicture}
    \end{minipage}
    \end{tabular}
    \caption{All possible directions of $w[i]$: straight, obtuse, acute.}
    \label{TTT_a3_w}
  \end{center}
\end{figure}

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a deterministic unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only finite conformations of length is $2n + 14$.
\end{theorem}

%Both $n_3$ and $n_4$ are occupied because $S[i] = t$ or $t^\times$.
%Suppose $w[i]$ were stabilized without any bond.
%Then, all neighbors of $n_3$ must be occupied as long as the bead at $n_3$ is not singular; otherwise, a bead at $n_3$ would be capable of forming a bond with $w[i]$.
%That is, $n_1$ is occupied.
%We can continue this argument in clockwise manner to fill the neighbors of $w[i]$ until a singular bead is encountered, if any.
%If needed, the same argument is applied starting from $n_4$ in counter-clockwise manner.
%Thus, all neighbors of $w[i]$ must be occupied but then $S[i+1] = \blacksquare$, contradiction. \qed

\begin{proof}
In this proof, we shall verify the claim that when the bead $w[i]$ is stabilized with $S[i] = b$ and $S[i+1] \neq \blacksquare$, if the circle of radius 2 centered at its predecessor $w[i-1]$ is free from the singular point, then $w[i]$ must form at least 2 bonds. 
Recall that the circle of radius 2 centered at the origin $O$, where the only candidate of singularity is, consists of 19 points including $O$. 
In order for the bead at $O$ to be singular, its $\alpha+1 = 4$ neighbors must be occupied. 
This means that there are at most 14 points where a bead find a singular bead within 2 points. 
Therefore, the claim, once proved, and the Tunnel Troll Theorem imply that all but at most 14 beads strictly decrease the binding capability. 
The binding capability of the seed is at most $3n$. 
Consequently, this theorem holds.

Now let us verify the claim. 
Suppose $w[i]$ were stabilized by just one bond. 
There are three cases to be considered as depicted in Fig.~\ref{TTT_a3_w}, depending on the relative position of $w[i]$ to $w[i-2]$ and $w[i-1]$. 
Since $S[i] = b$, at least one of the four neighbors of $w[i-1]$ must be empty. 
If $w_3$ is free, then $w[i-2]$ must have used up all of its hands; otherwise, $w[i]$ would be stabilized also at $n_3$ nondeterministically, a contradiction. 
Thus, $\alpha + 2 = 5$ neighbors of $w[i-2]$, that is, all of its neighbors, must be occupied. 
Hence, $n_5$ is occupied. 
(All the remaining arguments are based on this ``merry-go-round'' occupation. This works only if the circle of radius 2 around $w[i-1]$ is free from the singular bead). 
In the same way, all the neighbors of $n_3$ turned out to be occupied in the clockwise order, but eventually, we would encounter a neighbor that is adjacent to also the point where $w[i]$ is supposed to go. 
Thus, $n_3$ must be occupied (in the left and middle cases). 
In the left case, $n_4$ is symmetric to $n_3$, and hence, it must be occupied, too. 
Since $S[i] = b$, $n_1$ or $n_2$ must be free; assume $n_1$ is. 
Going clockwise around $n_1$ implies that $n_{-1}$ is occupied, but the bead at $n_1$ has a free hand and would cause a nondeterminism.

Let us focus on the remaining cases: middle and right. 
Suppose that among the 5 neighbors of $w[i]$ at which $w[i-1]$ is not, only one can be occupied. 
Otherwise, a bead without any hand is found at one of them, and staring from the point, merry-go-round occupies all the neighbors, but then $w[i+1]$ would lose its way to go. 
In the middle case, this means $n_0$ is free. 
Since $n_{-1}$ is free, so must be $n_2$. 
Repeating this, we get that both $n_4$ and $n_5$ are free, but then $w[i-2]$ would have a free hand and a free neighbor, attract $w[i]$, and cause a nondeterminism. 
Even in the right case, the points $n_1, n_0, n_2, n_4$ turn out to be free one after another likewise, but then $w[i-2]$ would have a free hand and neighbor, a contradiction. 
\qed

\end{proof}

%We consider cases which $w[i-1]$ is two distances away from the singular point.
%Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is, when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
%Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$.
%Suppose $w[i]$ were able to be stabilized with using one hand in each cases.
%
%% Then, $w[i]$ can be also stabilized at $n_3$. 
%
%\begin{paragraph}{Case of straight}
%\begin{itemize}
%\item[-] Case of $n_3$ is free\\
%The bead $w[i-2]$ must not have bonds because $w[i]$ is stabilized deterministically so that five neighbors of $w[i-2]$ must be occupied.
%Therefore $n_5$ must be occupied.
%We can continue this argument in clockwise manner to fill the neighbors of $n_3$ until $n_1$.
%However, two neighbors of $n_1$ must be free because they are $n_3$ and the point of $w[i]$, contradict.
%
%\item[-] Otherwise\\
%Both $n_3$ and $n_4$ are occupied.
%One of $n_1$ and $n_2$ must be free because $S[i] = b$.
%We assume $n_1$ is free, then all neighbors of $n_1$ must be occupied because $n_3$ is occupied.
%However, one of neighbors $n_1$ is a point which $w[i]$ is stabilized, contradict.
%\end{itemize}
%
%Therefore, this case contradict assumption.
%\end{paragraph}
%
%\begin{paragraph}{Case of obtuse}
%\begin{itemize}
%\item[-] Case of $n_3$ is free\\
%Any neighbors of $n_3$ have to be occupied but the point which $w[i]$ is stabilized is free. Thus $n_3$ has to be occupied.
%\item[-] Case of $n_4$ is free\\
%The point $n_2$ must be occupied because $n_4$ is free and $n_5$ must be occupied.
%Therefore $n_0$ is occupied.
%The bead $w[i]$ forms just one bond so that one of $n_0, n_3$ leave some hands or both of them do not leave any bonds.\\
%If $n_0$ has some hands, then $n_3$ does not have any hands so that $n_{-3}$ is occupied. 
%We can continue this argument in clockwise manner to fill the neighbors of $w[i]$.
%That is, $S[i] = \blacksquare$, contradict.
%
%If $n_3$ has some hands, then $n_0$ does not have any hands so that neighbors of $w[i]$ are occupied. In the same previous way, contradict.\\
%If both of $n_0, n_3$ do not have any hands, then both of $n_{-1}, n_{-3}$ are occupied. If one of $n_{-1}, n_{-3}$ has some hands, the other does not have any hands so that $n_{-2}$ is occupied. If both of $n_{-1}, n_{-3}$ do not have any hands, $n_{-2}$ has to be occupied and $n_{-2}$ has some hands, contradict.
%\item[-] Case of $n_2$ is free\\
%Any neighbors of $n_2$ have to be occupied so that $n_0$ is occupied. Any neighbors of $n_0$ except $n_2$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_2$ has to be occupied.
%\item[-] Case of $n_0$ is free\\
%Any neighbors of $n_0$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
%\end{itemize}
%\end{paragraph}
%
%
%
%\begin{paragraph}{Case of acute}
%\begin{itemize}
%\item[-] Case of $n_4$ is free\\
%The point $n_4$ and a point which is stabilized $w[i]$ are free so that  $w[i-2]$ has some bonds.
%However, $w[i]$ can be also stabilized $n_4$ in this case. Thus, $n_4$ has to be occupied.
%\item[-] Case of $n_2$ is free\\
%The point $n_0$ has to be occupied. The point $n_1$ has to be also occupied because five neighbors of $n_0$ must be occupied.
%We consider this case just like case of obtuse and that $n_4$ is free. Then if $w[i-2]$ binds $w[i]$, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. If $n_1$ binds $w[i]$, this case is same. Also if $n_1$ and $w[i-2]$ do not have any hand, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. Therefore, $w[i+1]$ cannot be provided.
%\item[-] Case of $n_0$ is free\\
%Any neighbors of $n_0$ have to be occupied so that $n_{1}$ is occupied. Any neighbors of $n_{1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
%\item[-] Case of $n_1$ is free\\
%Any neighbors of $n_1$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_1$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_1$ has to be occupied. 
%\end{itemize}
%
%Therefore, any situations contradict $S[i] = b$.
%\end{paragraph}\\
%
%
%Hence, $w[i]$ uses at least two bonds if $w[i-1]$ is at least two distances away the singular point.
%The size of $ \hexagon^2_0 = 19$ and this number contain the singular point, its successor and its three neighbors which is it binds to. Therefore 14 times  $w[i]$ can be stabilized by forming one bond.
%Therefore, a deterministic unary oritatami system at $\delta = 1$ and $\alpha = 3$ can yield only finite conformation of length is $2n + 14$.

\begin{figure}
  \begin{center}
  \begin{tabular}{c c}
  \begin{minipage}{0.48\hsize}
  \centering
    \begin{tikzpicture}
      	\fill (0,0) circle [radius=0.1];
        \fill[blue] (60:1) circle [radius=0.05];
         \fill[blue] (120:1) circle [radius=0.05];
        \fill (0:1) circle [radius=0.1];

      \draw[dashed] (180:2) -- (0:2);
      \draw[dashed, shift=(60:1)] (180:2.5) -- (0:1.5);
       \draw[dashed, shift=(60:2)] (180:3) -- (0:1);
      \draw[->, blue] (60:0.1) -- (60:0.9);
      \draw[->, blue, shift=(60:1)] (180:0.1) -- (180:0.9);

	\node[above] at (180:1) {$p_1$};
	\node[above] at (60:2) {$p_2$};
	\node[above,shift=(120:1)]at (60:1) {$p_3$};
	\node[above]at (0:1) {$p_4$};

	\draw (180:1) circle [radius=0.05];
	\draw (60:2) circle [radius=0.05];
	\draw[shift=(120:1)] (60:1) circle [radius=0.05];

	\node at (0:2.5) {$n$};
	\node[shift=(60:1)] at (0:2) {$n+1$};
	\node[shift=(60:2)] at (0:1.5) {$n+2$};

	\node[below] at (0:0) {$w[i-1]$};
	\node[above] at (60:1) {$w[i]$};
	\node[above] at (120:1) {$w[i+1]$};
	\end{tikzpicture}
    \end{minipage}

    \begin{minipage}{0.48\hsize}
    \begin{tikzpicture}
      	\fill (0,0) circle [radius=0.1];
        \fill[blue] (60:1) circle [radius=0.05];
         \fill[blue,shift=(0:1)] (60:1) circle [radius=0.05];
        \fill (0:1) circle [radius=0.1];

      \draw[dashed] (180:2) -- (0:2);
      \draw[dashed, shift=(60:1)] (180:2.5) -- (0:1.5);
       \draw[dashed, shift=(60:2)] (180:3) -- (0:1);
      \draw[->, blue] (60:0.1) -- (60:0.9);
      \draw[->, blue, shift=(60:1)] (0:0.1) -- (0:0.9);

	\node[above] at (180:1) {$p_1$};
	\node[above] at (60:2) {$p_2$};
	\node[above, shift=(120:1)]at (60:1) {$p_3$};
	\node[above]at (0:1) {$p_4$};

	\draw (180:1) circle [radius=0.05];
	\draw (60:2) circle [radius=0.05];
	\draw[shift=(120:1)] (60:1) circle [radius=0.05];

	\node at (0:2.5) {$n$};
	\node[shift=(60:1)] at (0:2) {$n+1$};
	\node[shift=(60:2)] at (0:1.5) {$n+2$};

	\node[below] at (0:0) {$w[i-1]$};
	\node[above] at (60:1) {$w[i]$};
	\node[above, shift=(0:1)] at (60:1) {$w[i+1]$};
	\end{tikzpicture}
    \end{minipage}
    \end{tabular}
    \caption{The bead $w[i]$  the first bead outside $\hexagon^n_0$: $w[i+1]$ elongated to west (left), $w[i+1]$ elongated to east (right)}
    \label{fig:ttt_d2}
  \end{center}
\end{figure}

\begin{theorem}[$\delta= 1, \alpha=2$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 2$. It can yield infinite structures but they are only zig-zag conformation in $w[27n^2 + 9n + 1]$.
\end{theorem}

\begin{proof}
 We assume $w[i]$ is the first bead outside $\hexagon^n_0$
 See Fig.~\ref{fig:ttt_d2}.
 If $w[i+1]$ elongated to west and $w[i+2]$ forms one bond, then $w[i+2]$ is stabilized at $p_3$; otherwise $w[i+2]$ forms two bonds.
 If $w[i+1]$ elongated to west and $w[i+2]$ is stabilized at $p_3$, $w[i+1..]$ form a zig-zag conformation.
  If $w[i+1]$ elongated to east and $w[i+2]$ forms one bond, then $w[i+2]$ is stabilized at $p_2$; otherwise $w[i+2]$ forms two bonds.
 If $w[i+1]$ elongated to east and $w[i+2]$ is stabilized at $p_2$, $w[i+1..]$ is zig-zag conformation.
 Thus, if $w[i]$ is the first bead outside $\hexagon^k_0$ for some $k \geq n$, $\#bc(C_i) > \#bc(C_{i+1})$ or $w[i+1..]$ form a zig-zag conformation.

 We assume $w[l]$ is the first bead outside $\hexagon^k_0$ and $w[m]$ is the first bead outside $\hexagon^{k+1}_0$ for some $k,l,m$ with $k<3n$. Let $s$ be a element of $S[l+1..m-1]$. If $s = b$, $s$ does not increase binding capability.
 If $s=t$, $s$ is a element of a subsequence $bt^+b$. This subsequence does not increase binding capability according to Theorem~\ref{TTT}. Thus, $\#bc(C_l) > \#bc(C_{m})$ or $w[l..]$ form a zig-zag conformation.

 The seed $\sigma$ provides at most $2n$ binding capability, that is, $\#bc(C_0) \leq 2n$.
 Thus,  if  $w[i]$ is the first bead outside $\hexagon^{3n}_0$, then this conformation has no hand without $w[i-1]$ and $w[i]$.
 Hence, $w[27n^2 + 9n + 1..]$ form a zig-zag conformation.
\end{proof}

%By Tunnel Troll Theorem, any tunnel sections which represented in $bbt^+$ or $bt^+bt^+$ consume binding capabilities. If the sequence $S$ is free from any subsequence of the form $bt^+bt^+$, then it can factorize as $S = u_1 u_2 u_3 \cdots$ for some $u_1 , u_2 , u_3 , \cdots \in \{b\} \cup bbt^+$. Assume the length of $\sigma$ is $n$, seed supplies at most $2n$ binding capabilities. Therefore formula \ref{TTT_only_bond} hold.
%
%\begin{eqnarray}
%  \exists i \in \mathbb{N} \quad s.t. \quad u_{i-1} , u_i , u_{i+1} , u_{i+2} , \cdots \in \{ b \}
%  \label{TTT_only_bond}
%\end{eqnarray}
%
%
%Let us represent $S$ as $S[i.i+1...] = v_i v_{i+1} v_{i+2} \cdots$ for some $v_i, v_{i+1}, v_{i+2}, \cdots \in \{ a, o\}$ where if $v_k$ is $a$, then $v_{k+1}$ is bound to $v_{k-1}$, if $v_k$ is $o$, then $v_{k+1}$ is NOT bound to $v_{k-1}$.
%
%
%Let us consider the case of $v_k$ is $o$. See Fig.\ref{TTT_case_of_o}. $w[i-1]$ consumes some binding capabilities because $v_{i-1}$ is $b$. If the number of $w[i-1]$'s bindings is one binding, then $w[i+1]$ has to be bound except $n_1$ or $n_2$ so that $w[i+1]$ must consumes two bindings except the case of $n_1$ and $n_2$ are occupied and $w[i]$ consumes at least one binding. If $n_1$ and $n_2$ are occupied, then $w[i-1]$'s bindings are inactive, that is, $w[i-1]$ consumes two binding capabilities. Therefore, this case consumes binding capabilities. If $w[i-1]$ dose Not have any bindings, then $w[i-1]$ already consumes two bindings. In addition, $w[i]$ and $w[i+1]$ consume at least one binding. Therefore this case consumes binding capabilities. Thus, the formula \ref{TTT_only_a} hold and according to the formula \ref{TTT_only_bond} and the formula \ref{TTT_only_a}, the formula \ref{TTT_structure} is hold. Thus, in this case, oritatami system can yield infinite structures but they are only zig-zag conformation.

%
%
%\begin{eqnarray}
%  \exists j \in \mathbb{N} \quad s.t. \quad u_j , u_{j+1} , u_{j+2} , \cdots \in \{ a \}
%  \label{TTT_only_a}
%\end{eqnarray}
%
%\begin{eqnarray}
%  | S | > \forall m \in \mathbb{N} \quad \to \quad \exists n \in \mathbb{N} \quad s.t. \quad S[n], S[n+1], \cdots \in \{ a \}
%  \label{TTT_structure}
%\end{eqnarray}
%
%\begin{figure}
%  \begin{center}
%    \begin{tikzpicture}
%      
%      \fill[transform canvas={shift=(-120:1)}](0,0) circle [radius=0.1];
%      
%      \fill[blue](0:0) circle [radius=0.05];
%      \fill[blue](0:1) circle [radius=0.05];
%      \fill[shift=(0:1), blue](40:1) circle [radius=0.05];
%
%      \draw[->, blue] (-120:1) edge[bend left] (-120:0.1);
%      \draw[->, blue] (0:0.1) -- (0:0.9);
%      \draw[->,transform canvas={shift=(0:1.2)}, blue] (50:0) -- (47:0.8);
%      
%      \draw[shift=(60:1)](-0.05,-0.05) rectangle (0.05,0.05);
%      \draw[shift=(-60:1)](-0.05,-0.05) rectangle (0.05,0.05);
%      
%      \node[below] at (-120:1) {$w[i-2]$};
%      \node[above] at (0:0) {$w[i-1]$};
%      \node[below] at (0:1) {$w[i]$};
%      \node[right, shift=(0:1)] at (40:1) {$w[i+1]$};
%
%      \node[above] at (60:1) {$n_1$};
%      \node[below] at (-60:1) {$n_2$};
%        
%    \end{tikzpicture}
%    \caption{Case of $S[i]$}
%    \label{TTT_case_of_o}
%  \end{center}
%\end{figure}
%